---
title: "Maximum-likelihood tree for examining biomass effects caused by the 8 strain subset"
author: " Maria Alejandra Gil Polo - mag6678@psu.edu, Elizabeth Leslie Paillan - elpailla@ncsu.edu, Liana Burghardt - liana.burghardt@psu.edu"
date: "2026-04-27"
output: html_document
---

This file cointains the information that has been used to create the phylogeny of the 117 strains used for the nodule branching experiment and it also includes information about the mean raw biomass, mean raw proportion of branched nodules, and their origin, which means from which of the three alfalfa hosts were they originally isolated from. The tree used was generated using the maximum-likelihood plot. For more information about how this tree was created see Guha et al., In preparation. 


```{r libraries,  echo = TRUE, warning=FALSE}

library(tidyverse)
library(ggpubr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(ggtree)
library(vegan)
library(corrplot)
library(patchwork)
library(ggnewscale)
library(ape)
library(readxl)
library(ggtreeExtra)
library(colorRamps)

ss_df <- read.csv("/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/ss_df.csv", colClasses = c("Strain" = "character")) ## Make sure the name of the strains are getting imported as characters, otherwise they will transform to numerical values, for example Strain 001 will become 1. 

#View(ss_df)

## I am only going to select the traits I am interested in which are Shootweight_g and prob_lobe
ss_df_subset <- ss_df %>% select(Strain, Variety, Shootweight_g, prop_lobe)

ss_df_subset_mean <- ss_df_subset %>%
  group_by(Strain, Variety) %>%
  summarise(mean_shoot = mean(Shootweight_g),
            mean_prop_lobe = mean(prop_lobe))

### I will modify the dataframe now so each column is an alfalfa variety

ss_df_shoot_mean_pivot <- ss_df_subset_mean %>%
  pivot_wider(
    names_from = Variety,   # Variety values become column names
    values_from = c(mean_shoot, mean_prop_lobe),   # Biomass and prop of lobed nodules values fill the cells, 
    names_glue = "{Variety}_{.value}"
  )

#View(ss_df_shoot_mean_pivot)

```

## For the actual tree

The aesthetics of the tree were mostly arranged in the itols app. However, I created the metadata to be imported to itols in the code bellow

```{r}
### These files were done by Sohini Guha. 

tree<-read.tree("../Data/AVT_single_copy_core_concat.fasta.treefile")

# 2. Read your annotation file from iTOL
# Replace the path with wherever you saved Annotation_iTOL_final.txt
meta_raw <- read.delim("../Data/Annotation_iTOL_final.txt", sep = "\t", header = TRUE)
meta_raw ## This is the meta_raw used in Sohini's paper: Morphology and origin

###  I am trying to work with the codes that I used 
Codes <- read_excel("/Users/mariaalejandragilpolo/Downloads/Strain_codes_ForTree.xlsx") #Has all the 117 strains and information on mucoidy and pairs. I created this because I am working with a different format of strain names. 
### I will filter out the Morphology and Pair that I will not use for this analysis 

Codes <- Codes %>% select(Strain, Strain_Code)


## Rename Strain as Strain_Code

ss_df_shoot_mean_pivot_renamed <- ss_df_shoot_mean_pivot %>% dplyr::rename(Strain_Code = Strain) 
#View(ss_df_shoot_mean_pivot_renamed)
ss_df_pivot_withCodes <- left_join(Codes, ss_df_shoot_mean_pivot_renamed, by = "Strain_Code") 
#View(ss_df_pivot_withCodes)
### I will merge that file with meta_raw created by Sohini that has the "Variety_source" that represents where is the rhizobia isolated from 
#View(meta_raw)
meta_tree <- left_join(meta_raw %>% select(Strain, Source_variety), ss_df_pivot_withCodes, by = "Strain")
#View(meta_tree) # This dataframe contains the information I need: Strain, Strain_Code (for the tips), the shoot biomass mean and the proportion of brahcing
```

### Create the metadata for the tree

The idea is that the metadata has shape plots for aboveground biomass in circles and proportion branched in squares

```{r}
### Start with the labels 


#### labels_df for labels. 
labels_df <- data.frame(
  tip = meta_tree$Strain,
  new_label = meta_tree$Strain_Code
)
labels_df <- labels_df[!is.na(labels_df$tip) & !is.na(labels_df$new_label), ]
labels_df

#View(meta_tree)

### Create a dataframe with the datapoints I care about: The aboveground biomass data is in Shoot_biomass

### Now I will write them in a format for itols 

cat("LABELS\nSEPARATOR TAB\n\nDATA\n",
    file = "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/labels.txt")
write.table(labels_df, "../Tables/labels.txt",
            sep = "\t", append = TRUE, row.names = FALSE,
            col.names = FALSE, quote = FALSE)
```

###Start with shoot_biomass


```{r}
# Generate iTOL DATASET_SYMBOL file: Shape plot
# The idea is to have three colored circles per strain, one per host (O, V, S) that will have different sizes depeding on how big the raw biomass is.
# The circle size is raw biomass on that host.
# Circle color = host (O = red, V = pinkish, S = blue).## I am using the same colors that Liz used
# Input:  meta_tree with columns Strain, O, V, S
# Output: itol_annotations/biomass_symbols.txt

host_colors <- c(
  "O_mean_shoot" = "#471323",   
  "V_mean_shoot" = "#FCB0B3",  
  "S_mean_shoot" = "#ADB9E3"   
)
 
# Create the file

#dir.create("/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/itol_annotations", showWarnings = FALSE)
 
df <- meta_tree   # I am using df to replace meta_tree 
 
## Build the table that will be the input in itols. 
# DATASET_EXTERNALSHAPE expects: ID  value_field1  value_field2  value_field3
# Sizes are scaled automatically across all values in the dataset.
 
shape_df <- df[, c("Strain", "O_mean_shoot", "V_mean_shoot", "S_mean_shoot")]
shape_df
shape_df <- shape_df[!is.na(shape_df$Strain) & nzchar(shape_df$Strain), ]
shape_df

outfile <- "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/itol_annotations/biomass_externalshape.txt"

 ## Writing the file. Dataset_externalshape is important to keep them separate 
cat(
  "DATASET_EXTERNALSHAPE\n",
  "SEPARATOR TAB\n",
  "DATASET_LABEL\tShoot biomass\n",
  "COLOR\t#000000\n",
 
  # One column per host, with column headers
  "FIELD_LABELS\tO_mean_shoot\tV_mean_shoot\tS_mean_shoot\n",
  "FIELD_COLORS\t", host_colors["O_mean_shoot"], "\t",
                    host_colors["V_mean_shoot"], "\t",
                    host_colors["S_mean_shoot"], "\n",
 
  # Shape per column: 2 = filled circle (per iTOL docs)
  "SHAPE_TYPE\t2\t2\t2\n",
 
  "SHAPE_SCALING\t1\n",        # scale shapes by value (the whole point)
  "SHOW_VALUES\t0\n",          
 
  # Visual spacing
  "SHAPE_SPACING\t10\n",       # gap between columns
  "HORIZONTAL_GRID\t1\n",
  "VERTICAL_GRID\t1\n",
 
  # Legend
  "LEGEND_TITLE\tBiomass by host\n",
  "LEGEND_SHAPES\t2\t2\t2\n",
  "LEGEND_COLORS\t", host_colors["O_mean_shoot"], "\t",
                     host_colors["V_mean_shoot"], "\t",
                     host_colors["S_mean_shoot"], "\n",
  "LEGEND_LABELS\tO_mean_shoot\tV_mean_shoot\tS_mean_shoot\n",
 
  "\nDATA\n",
  sep = "",
  file = outfile
)


write.table(shape_df, outfile,
            sep = "\t", append = TRUE,
            row.names = FALSE, col.names = FALSE,
            quote = FALSE, na = "X")
 
cat("Wrote:", outfile, "\n")
cat("Drag it onto your tree in iTOL.\n")
cat("If circles are too big/small, adjust via the gear icon next to the dataset.\n")
 

```

### Proportion branched

```{r}
### Start with proportion branched

# Generate iTOL DATASET_SYMBOL file: Shape plot
# The idea is to have three colored squares per strain, one per host (O, V, S) that will have different sizes depeding on how big the raw proportion branched is is.
# The square size is raw biomass on that host.
# Square color = host (O = red, V = pinkish, S = blue).## I am using the same colors that Liz used
# Input:  meta_tree with columns Strain, O, V, S (with the _mean_prop_lobe)
# Output: itol_annotations/biomass_symbols.txt

host_colors <- c(
  "O_mean_prop_lobe" = "#471323",   
  "V_mean_prop_lobe" = "#FCB0B3",  
  "S_mean_prop_lobe" = "#ADB9E3"   
)
 
# Create the file

#dir.create("/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/", showWarnings = FALSE)
 
df <- meta_tree   # I am using df to replace meta_tree 
 
## Build the table that will be the input in itols. 
# DATASET_EXTERNALSHAPE expects: ID  value_field1  value_field2  value_field3
# Sizes are scaled automatically across all values in the dataset.
 
shape_df <- df[, c("Strain", "O_mean_prop_lobe", "V_mean_prop_lobe", "S_mean_prop_lobe")]
shape_df
shape_df <- shape_df[!is.na(shape_df$Strain) & nzchar(shape_df$Strain), ]
shape_df

outfile <- "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/itol_annotations/Proportion_Branched_externalshape.txt"

 ## Writing the file. Dataset_externalshape is important to keep them separate 
cat(
  "DATASET_EXTERNALSHAPE\n",
  "SEPARATOR TAB\n",
  "DATASET_LABEL\tProportion of branched nodules\n",
  "COLOR\t#000000\n",
 
  # One column per host, with column headers
  "FIELD_LABELS\tO_mean_prop_lobe\tV_mean_prop_lobe\tS_mean_prop_lobe\n",
  "FIELD_COLORS\t", host_colors["O_mean_prop_lobe"], "\t",
                    host_colors["V_mean_prop_lobe"], "\t",
                    host_colors["S_mean_prop_lobe"], "\n",
 
  # Shape per column: 1 = filled square (per iTOL docs)
  "SHAPE_TYPE\t1\t1\t1\n",
 
  "SHAPE_SCALING\t1\n",        # scale shapes by value (the whole point)
  "SHOW_VALUES\t0\n",          
 
  # Visual spacing
  "SHAPE_SPACING\t10\n",       # gap between columns
  "HORIZONTAL_GRID\t1\n",
  "VERTICAL_GRID\t1\n",
 
  # Legend
  "LEGEND_TITLE\tProportion Branched by host\n",
  "LEGEND_SHAPES\t1\t1\t1\n",
  "LEGEND_COLORS\t", host_colors["O_mean_prop_lobe"], "\t",
                     host_colors["V_mean_prop_lobe"], "\t",
                     host_colors["S_mean_prop_lobe"], "\n",
  "LEGEND_LABELS\tO_mean_prop_lobe\tV_mean_prop_lobe\tS_mean_prop_lobe\n",
 
  "\nDATA\n",
  sep = "",
  file = outfile
)


write.table(shape_df, outfile,
            sep = "\t", append = TRUE,
            row.names = FALSE, col.names = FALSE,
            quote = FALSE, na = "X")
 
cat("Wrote:", outfile, "\n")
cat("Drag it onto your tree in iTOL.\n")
cat("If squares are too big/small, adjust via the gear icon next to the dataset.\n")
 

```
If I add up the host O, S, V
2 strains originated in O, three in V and 3 in S. 

001 originated in O
070 originated in V
077 originated in V
012 originated in O
117 originated in V
059 originated in S 
112 originated in S 
044 originated in S 

```{r}

# ============================================================
# Edit these three colors to change Morphology appearance
origin_colors <- c("S" = "#FA9169", "O" = "#6EC3AA", "V" = "#91A5CD")

origin_df <- data.frame(
  tip   = meta_tree$Strain,
  color = origin_colors[as.character(meta_tree$Source_variety)],
  label = meta_tree$Source_variety
)
origin_df <- origin_df[!is.na(origin_df$color), ]

cat(
  "DATASET_COLORSTRIP\n",
  "SEPARATOR TAB\n",
  "DATASET_LABEL\tSource Variety\n",
  "COLOR\t#000000\n",
  "STRIP_WIDTH\t25\n",
  "LEGEND_TITLE\tSource Variety\n",
  "LEGEND_SHAPES\t2\t2\t2\n",
  "LEGEND_COLORS\t", origin_colors["S"], "\t", origin_colors["V"], "\t", origin_colors["O"], "\n",
  "LEGEND_LABELS\tS\tV\tO\n",
  "\nDATA\n",
  sep = "",
  file = "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/itol_annotations/origin_colorstrip.txt"
)
write.table(origin_df, "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/itol_annotations/origin_colorstrip.txt",
            sep = "\t", append = TRUE, row.names = FALSE,
            col.names = FALSE, quote = FALSE)


```

