---
title: "Maximum-likelihood tree for examining biomass effects caused by the 8 strain subset"
author: " Maria Alejandra Gil Polo - mag6678@psu.edu, Elizabeth Leslie Paillan - elpailla@ncsu.edu, Liana Burghardt - liana.burghardt@psu.edu"
date: "2026-04-27"
output: html_document
---
This RMD contains code for generating the biomass maximum-likelihood plot we used to show how different the 8 strains we used span across our 117 strain collection, and to show how 

## Mariale trying to figure out how to better represent the span of shoot biomass 

```{r libraries,  echo = TRUE, warning=FALSE}

library(tidyverse)
library(ggpubr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(ggtree)
library(vegan)
library(corrplot)
library(patchwork)
library(ggnewscale)
library(ape)
library(readxl)
library(ggtreeExtra)
library(colorRamps)


ss_df<-data.frame(fread("../data/MasterSpreadsheet_AVT.csv"))
# Removing erroneous columns in the original data frame, this includes collection notes
ss_df <- ss_df[, !colnames(ss_df) %in% "NOTES"]

# Adding total nodule count and the proportion of branched nodules
ss_df<-ss_df %>%
  mutate(Nod_count=Nod_lobe+Nod_unlobed)%>% #Counting total observed nodules
  mutate(prop_lobe=Nod_lobe/Nod_count) %>% #Creating the proportion of branched nodules to nodule count
  mutate(Biomass_g=Rootweight_g+Shootweight_g) #Coutning total host mass biomass


## I am going to filter out MIX
ss_df <- ss_df %>% 
  filter(!Strain %in% c("MIX", "NIG", "MC2", "EXC", "UNI")) %>% 
    filter(!is.na(Strain), Strain != "")



## I am only going to select the things I am interested in 
ss_df_shoot <- ss_df %>% select(Strain, Variety, Shootweight_g)


#ss_df_shoot <- ss_df %>% select(Strain, Variety, Shootweight_g) ## According to her code the mean of shoot biomass is named mean_abiomass_g
ss_df_shoot_mean <- ss_df_shoot %>%
  group_by(Strain, Variety) %>%
  summarise(shoot_ave = mean(Shootweight_g, na.rm = TRUE))

View(ss_df_shoot_mean)

Shoot_biomass <- ss_df_shoot_mean %>%
  pivot_wider(
    names_from = Variety,   # Variety values become column names
    values_from = shoot_ave   # Biomass values fill the cells
  )

Shoot_biomass

```

## For the actual tree

The aesthetics of the tree were mostly arranged in the itols app. However, I created the metadata to be imported to itols in the code bellow
```{r}
### These files were done by Sohini Guha. 
tree<-read.tree("../Data/AVT_single_copy_core_concat.fasta.treefile")


# 2. Read your annotation file from iTOL
# Replace the path with wherever you saved Annotation_iTOL_final.txt
meta_raw <- read.delim("../Data/Annotation_iTOL_final.txt", sep = "\t", header = TRUE)
meta_raw ## This is the meta_raw used in Sohini's paper: Morphology and origin

###  I am trying to work with the codes that I used 
Codes <- read_excel("/Users/mariaalejandragilpolo/Downloads/Strain_codes_ForTree.xlsx") #Has all the 117 strains

## Rename Strain as Strain_Code

Shoot_biomass_edit <- Shoot_biomass %>% dplyr::rename(Strain_Code = Strain) 

Shoot_biomass_withCodes <- left_join(Codes, Shoot_biomass_edit, by = "Strain_Code") 


### Start with the labels 


#### labels_df for labels. 
labels_df <- data.frame(
  tip = meta_tree$Strain,
  new_label = meta_tree$Strain_Code
)
labels_df <- labels_df[!is.na(labels_df$tip) & !is.na(labels_df$new_label), ]
labels_df


### Create a dataframe with the datapoints I care about: The aboveground biomass data is in Shoot_biomass

meta_tree <- meta_raw %>%
  left_join(
    Shoot_biomass_withCodes %>%
      select(Strain, Strain_Code, O, V, S), 
    by = "Strain"
  ) %>%
  select(-Morphology, -Source_variety)

#View(meta_tree) Looking good

### Now I will write them in a format for itols 

cat("LABELS\nSEPARATOR TAB\n\nDATA\n",
    file = "../Tables/labels.txt")
write.table(labels_df, "../Tables/labels.txt",
            sep = "\t", append = TRUE, row.names = FALSE,
            col.names = FALSE, quote = FALSE)


# Generate iTOL DATASET_SYMBOL file: Shape plot
# The idea is to have three colored circles per strain, one per host (O, V, S) that will have different sizes depeding on how big the raw biomass is.
# The circle size is raw biomass on that host.
# Circle color = host (O = red, V = pinkish, S = blue).## I am using the same colors that Liz used
# Input:  meta_tree with columns Strain, O, V, S
# Output: itol_annotations/biomass_symbols.txt

host_colors <- c(
  "O" = "#471323",   
  "V" = "#FCB0B3",  
  "S" = "#ADB9E3"   
)
 
# Create the file

 
dir.create("itol_annotations", showWarnings = FALSE)
 
df <- meta_tree   # I am using df to replace meta_tree 
 

## Build the table that will be the input in itols. 
# DATASET_EXTERNALSHAPE expects: ID  value_field1  value_field2  value_field3
# Sizes are scaled automatically across all values in the dataset.
 
shape_df <- df[, c("Strain", "O", "V", "S")]
shape_df
shape_df <- shape_df[!is.na(shape_df$Strain) & nzchar(shape_df$Strain), ]
 shape_df

 outfile <- "itol_annotations/biomass_externalshape.txt"

 ## Writing the file. Dataset_externalshape is important to keep them separate 
cat(
  "DATASET_EXTERNALSHAPE\n",
  "SEPARATOR TAB\n",
  "DATASET_LABEL\tShoot biomass\n",
  "COLOR\t#000000\n",
 
  # One column per host, with column headers
  "FIELD_LABELS\tO\tV\tS\n",
  "FIELD_COLORS\t", host_colors["O"], "\t",
                    host_colors["V"], "\t",
                    host_colors["S"], "\n",
 
  # Shape per column: 2 = filled circle (per iTOL docs)
  "FIELD_SHAPES\t2\t2\t2\n",
 
  "SHAPE_SCALING\t1\n",        # scale shapes by value (the whole point)
  "SHOW_VALUES\t0\n",          
 
  # Visual spacing
  "SHAPE_SPACING\t10\n",       # gap between columns
  "HORIZONTAL_GRID\t1\n",
  "VERTICAL_GRID\t1\n",
 
  # Legend
  "LEGEND_TITLE\tBiomass by host\n",
  "LEGEND_SHAPES\t2\t2\t2\n",
  "LEGEND_COLORS\t", host_colors["O"], "\t",
                     host_colors["V"], "\t",
                     host_colors["S"], "\n",
  "LEGEND_LABELS\tO\tV\tS\n",
 
  "\nDATA\n",
  sep = "",
  file = outfile
)
 
write.table(shape_df, outfile,
            sep = "\t", append = TRUE,
            row.names = FALSE, col.names = FALSE,
            quote = FALSE, na = "X")
 
cat("Wrote:", outfile, "\n")
cat("Drag it onto your tree in iTOL.\n")
cat("If circles are too big/small, adjust via the gear icon next to the dataset.\n")
 


```

