---
title: "Nodule branching, size, and symbiosis outcomes are shaped by natural genetic variation in rhizobial bacteria and alfalfa hosts "
author: "Elizabeth Leslie Paillan - elpailla@ncsu.edu, Maria Alejandra Gil Polo - mag6678@psu.edu, Liana Burghardt - liana.burghardt@psu.edu"
date: "2025-10-25"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

Nodule traits being examined: Proportion Lobed (prop_lobe= \# of lobed nodule/ total \# nodule sample) a lower proportion means fewer lobed nodules, and a higher proportion means larger number of nodules. Nodule Prevalence (nod_prev=total \# nodule sample), the term origniates from the avt_field experiement where I was dividing the number of nodules found by plant, that is not needed since its was one plant (mostly), for this experiment its just the total nodule number.

#Libraries Used

```{r Library setup, echo = TRUE, warning=FALSE}

library(tidyverse); packageVersion("tidyverse")
library(BSDA); packageVersion("BSDA")
library(patchwork); packageVersion("patchwork")
library(svglite); packageVersion("svglite")
library(performance); packageVersion("performance")
library(data.table); packageVersion("data.table")
set.seed(100)

```

```{r GGplot theme, include=FALSE}
#Below is the general theme used for all ggplot2 figures
plot_theme <- function(base_size = 18) {
  theme_bw(base_size = base_size) +
    theme(legend.title = element_text(face = "bold"),
      legend.text  = element_text(face = "bold"),
      axis.title   = element_text(face = "bold"),
      strip.text   = element_text(face = "bold"),
      panel.grid   = element_blank(),
      panel.background = element_blank(),
      axis.line    = element_line(linewidth = 1.05, color = "black"))
}

#Changing host variety names to their full names & add designated color scheme
host_labels <- c(O = "Onedia", S = "SW402", V = "Vernal")
host_colors <- c(O = "#471323", S = "#ADB9E3", V = "#FCB0B3")

#same thing while nodule branch number
nbranch_labels <- c("1", "2", "3-4", "5+")
nbranch_colors <- c("1" = "#273E47",
                    "2" = "#BAA5FF",
                    "3-4" = "#6E9887",
                    "5+" = "#A5402D")
nbranch_shapes <- c("1" = 16, "2" = 15, "3-4" = 17, "5+" = 18)

#and strain ID
strain_labels <- c("001"=1, "012"=2, "044"=3, "059"=4, "070"=5, "077"=6, "112"=7, "117"=8)
strain_shapes <- c("001"=15, "012"=16, "044"=17, "059"=18, "070"=8, "077"=4, "112"=25, "117"=3)

```

#Functions written for the analysis:

```{r Function: Anova on linear models -> pulling the summary and r2, echo=TRUE}

### This function runs a general anova, and spit out the summary() table and export it

lm_run<- function(model, model_name, save_path=getwd()) {
  anova_output<- anova(model)
  model_summary <- summary(model)
  results_list<- list()
  
  results_list[[paste0(model_name, "_anova")]] <- anova_output
  results_list[[paste0(model_name, "_summary")]] <- model_summary
  
  anova_df<- as.data.frame(anova_output)
  summary_df<- as.data.frame(model_summary$coefficients)
  
  return(results_list)
}
```

```{r Function: Detecting and Listing Outliers, }
### Maria: Why are you dropping these outliers? Have you tried running the code first with the "outliers" to see how the results look like?. 

#This function is used to detect outliers, create a list of the identified outliers.
detectview_outlier <- function(data, variable) {
  outliers <- boxplot(data[[variable]], plot = FALSE)$out #Generates a barplot to ID outliers
  outlier_data <- data[data[[variable]] %in% outliers, ]
  
  outlier_plot <- ggplot(data, aes(x = 1, y = !!rlang::sym(variable))) + 
    geom_boxplot() + 
    geom_point(data = outlier_data, aes(x = 1, y = !!rlang::sym(variable)), color = "red") + #Makes the outlier more visible
    labs(x = "", y = variable, title = paste("Boxplot of", variable, "outliers"))
  
  return(list(plot = outlier_plot, outlier_data = outlier_data)) #Provides detail about the identified outliers
}
```

#Importing Data

```{r Data Import of Single Inoculation (ss_df), results = HIDE, warning = HIDE, collapse=TRUE}
#Import and tidying of Data
setwd("~/GitHub/nodule_branching_morphology/Analysis/") #Maria: This gives me error and does not seem to be relevant. 

getwd()

#Setting up plot and directories to be easily found:
plot_dir<-("../Figures/")
table_dir<-("../Tables/")

ss_df<-data.frame(fread("../data/MasterSpreadsheet_AVT.csv"))

# Removing erroneous columns in the original data frame, this includes collection notes
ss_df <- ss_df[, !colnames(ss_df) %in% "NOTES"]

# Adding total nodule count and the proportion of branched nodules
ss_df<-ss_df %>%
  mutate(Nod_count=Nod_lobe+Nod_unlobed)%>% #Counting total observed nodules
  mutate(prop_lobe=Nod_lobe/Nod_count) %>% #Creating the proportion of branched nodules to nodule count
  filter_at(vars(Nod_count:prop_lobe), all_vars(!is.infinite(.)))%>% #Checking for any infinite values
  mutate(Biomass_g=Rootweight_g+Shootweight_g)%>% #Counting total host mass biomass
  na.omit(ss_df) #Removing any NAs ## Maria: This code is eliminating the rows across all the datadram that have NA. While I understand that for the correlations you want to drop the NAs, I think that not all of the traits are being used for the correlations, so you should get rid of only the NAs for the important columns, not across the dataframe that could eliminate important samples if they happen to have a NA for any trait. I noticed that this code might be a bit problematic because it eliminated 10 out of the 12 Uninoculated samples. 

# Separate data frame that is the mean of every Isolate-Host variety combination across all replicates
Mean_lobe<-ss_df %>% group_by(Strain,Variety,Source_variety,Chloro_avg)%>% summarize(lobe_mean=mean(prop_lobe, na.rm=TRUE))

# Using the detect outlier function to isolate them, no major outliers found in biomass
Biom_outlier<-detectview_outlier(ss_df,"Biomass_g")
Lobe_outlier<-detectview_outlier(ss_df,"prop_lobe") #Removing plants that have a proportion of 1 meaning all nodules we're branched -> most likely a counting error # Maria - why do you think this is a counting error? It is possible that all the noduels of the plants have branched nodules

# Get the outlier data by XYZ_outlier$outlier_data  ## Maria: XYZ_outlier does not exist in your data analysis, please make sure you are removing confusing explanation parts. 
Lobe_outlier$outlier_data
# Removing found outliers
Prop_lobe.outliers<-c("110_O_a","EXC_O_d")  ## Maria: I am not sure why are you still maintaining EXC. Shouldn't the analysis be done only with the 117 strains and not the commercial strains. If not this should be stated in the manuscript

ss_df<-ss_df %>%
  filter(!(Label %in% Prop_lobe.outliers))

```

Maria: You should drop things that are not important for your analysis, for example Strain MIX", "NIG", "MC2", "EXC", "UNI". I have not seeing anything in your manuscript that suggests that you are using MIX NIG MC2 EXC and UNI as your samples. I think you are only focusing on the 117 strains in the draft. Filter them out before you analyze the data. 


```{r Data Import of CFU experiment (cfu_df), message=FALSE, warning=FALSE, page.print=FALSE}
#Loading in the cfu dataset
cfu_df<-data.frame(fread("../data/AVT_CFU_count.csv"))

#Removing erroneous columns from the data frame like notes from the harvest, and renames one column to be easier to call into R
cfu_df <- cfu_df %>%
  mutate(Comments=NULL)%>%
  mutate(dilution.counted=NULL)%>%
  rename(rhz_released = Lianas.rhizobia.released.per.nodule.calculation)%>% 
#Since we had a couple of nodules that had higher than 5 branches we combined them into a 5+ category
  mutate(Nodule.lobes = ifelse(Nodule.lobes == "more than 5", "5+", 
                        ifelse(Nodule.lobes == "More than 5", "5+",
                        ifelse(Nodule.lobes == "5", "5+",
#Few nodules had 3-4 branches so we combined them into a 3-4 category
                        ifelse(Nodule.lobes %in% c("3", "4"), "3-4", Nodule.lobes)))))%>%
#Making sure the branch coulmn is considered a factor
  mutate(Nodule.lobes = as.factor(Nodule.lobes))

#Extracting sample data from the sample-ID
cfu_df<-separate(cfu_df, Plant, into = c("Strain","Host","rep"), sep="\\.")

#Ensuring intended variables are numeric and factor as intended
cfu_df$rhz_released<-as.numeric(cfu_df$rhz_released)
cfu_df$Nodule.area<-as.numeric(cfu_df$Nodule.area)
cfu_df$spot.avg <- as.numeric(gsub("[^0-9.]", "", cfu_df$spot.avg))


#Detecting outliers in the CFU data
RhzReleased_outlier<-detectview_outlier(cfu_df,"rhz_released")
Nodule.area_outlier<-detectview_outlier(cfu_df,"Nodule.area")

#Removing detected outliers
RhzReleased.outliers<-c("070.O.b.5")
Nodule.area.outliers<-c("044.V.b.4")

cfu_df<-cfu_df %>%
#Only removing the largest outlier
  filter(!(Ind.Label %in% RhzReleased.outliers))%>%
#Removing size that was likely an error during processing
  filter(!(Ind.Label %in% Nodule.area.outliers))%>%
  na.omit(cfu_df)

#Removing any NA's introduced by adding a log transformation to the colony counts
cfu_df[is.infinite(log10(cfu_df$rhz_released)) | 
         is.na(log10(cfu_df$rhz_released)), ]
cfu_df <- subset(cfu_df, rhz_released > 0 & !is.na(rhz_released))

```

```{r Strain Reaction norm df (str.host_mean_df), collapse=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Another dataframe to create a reaction norm like plot
#Taking the mean across reps, strain, and variety to reduce noise
str.host_mean_df <- ss_df %>%
  group_by(Strain,Variety)%>%
  summarise(mean_prop_lobe=mean(prop_lobe,na.rm=TRUE), 
            mean_chloro_avg=mean(Chloro_avg, na.rm=TRUE),
            mean_biomass_g=mean(Biomass_g, na.rm=TRUE),
            mean_nod_num=mean(Nod_count, na.rm=TRUE),
            mean_agbiomass_g=mean(Shootweight_g, na.rm=TRUE))
```

#Statistics used 
##Rhizboial focused traits 
```{r Rhizobial Trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova examining Nodule area response predicted by branch #, host variety,and isolate ID
nodarea<-lm(log10(Nodule.area)~Nodule.lobes*Host*Strain+rep, data=cfu_df)
nodarea.results<-lm_run(nodarea, model_name="nodarea")
aov.nodarea<-anova(nodarea)
write.csv(aov.nodarea, file=paste(table_dir,"Nodule-Area-response_anova.csv"))
aov.nodarea
aov.nodarea[["Sum Sq"]] / sum(aov.nodarea[["Sum Sq"]])

#Model examining Amount of viable rhizobia released focused on number of nodule branches as a predictor
rhz_released<-lm(log10(rhz_released)~as.factor(Nodule.lobes)*Host+as.factor(Strain)+rep, data=cfu_df)
rhz_released.results<-lm_run(rhz_released, "rhzreleased")
aov.rhzreleased<- anova(rhz_released)
write.csv(aov.rhzreleased, file=paste(table_dir,"RhizobiaReleased-response_anova.csv"))
aov.rhzreleased
aov.rhzreleased[["Sum Sq"]] / sum(aov.rhzreleased[["Sum Sq"]])


# Model examining amount of rhizobia released focused on nodule surface area as a predictor 
rhztestarea<-lm(log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep, data=cfu_df)
rhztestarea.results<-lm_run(rhztestarea,"rhzarea")
aov.rhztestarea<- anova(rhztestarea)
write.csv(aov.rhztestarea, file=paste(table_dir,"Nodulearea-by-surfacearea_anova.csv"))
aov.rhztestarea
aov.rhztestarea[["Sum Sq"]] / sum(aov.rhztestarea[["Sum Sq"]])


```

##Statistical Analysis of the Host plant traits
```{r host plant trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova is being ran on models associated with host traits 
#Modeling the mean nodule number predicted by proportion of branched nodules and host variety
nodnumber<-lm(mean_nod_num~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
nodnum.results<-lm_run(nodnumber, "nodnum")
aov.nodnum<-anova(nodnumber)
write.csv(aov.nodnum, file=paste(table_dir,"Nodulenumber-response_anova.csv"))
aov.nodnum
aov.nodnum[["Sum Sq"]] / sum(aov.nodnum[["Sum Sq"]])


#Modeling the mean host plant biomass predicted by the mean proportion of branched nodules and host variety
biomass<-lm(mean_biomass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
biomass.results<-lm_run(biomass, "biomass")
aov.biomass<-anova(biomass)
write.csv(aov.biomass, file=paste(table_dir,"Biomass-response_anova.csv"))
aov.biomass
aov.biomass[["Sum Sq"]] / sum(aov.biomass[["Sum Sq"]])


propbr<- lm(prop_lobe ~ Variety*Strain+Rep,data=ss_df)
propbr.results<- lm_run(propbr,"meanbranch")
aov.propbr<-anova(propbr)
write.csv(aov.propbr, file=paste(table_dir,"proplobe-response_anova.csv"))
aov.propbr
aov.propbr[["Sum Sq"]] / sum(aov.propbr[["Sum Sq"]])




```

#Figures

##Figure 2
```{r Figure 2, echo=FALSE, warning=FALSE, message=TRUE}
#Single plot exporting has been "#" out to make this knit, if an export is needed removing the "#" from the svg and dev.off() functions below.

#svg(filename=file.path(plot_dir,"Figure_2_surfacearea.svg"), width=3, height=3)
surface_area_plot<- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=Nodule.area, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black")+
  labs(x="", y="Nodule Surface Area(cm2)", shape="Rhizobial\nIsolates", color="Number of\nNodule\nBranches")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  plot_theme()
#dev.off()

#svg(filename=file.path(plot_dir,"Figure_2_released_rhizobia_branch.svg"), width=3, height=3)
rhzreleasedbr_plot <- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=rhz_released, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black")+
  scale_y_log10()+
  labs(x= "Number of Nodule Branches", y="CFU/Nodule")+
  labs(shape="Rhizobial\nIsolates")+
  labs(color="Host Variety")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  plot_theme()+
  theme(legend.position = "none")
  
#dev.off()

#svg(filename=file.path(plot_dir,"Figure_2_released_rhizobia_branch_ver5.svg"), width=3, height=3)
rhzreleased_plot<-
  ggplot(cfu_df, aes(x = Nodule.area, y = rhz_released, color = Nodule.lobes)) +
  geom_point(size=3)+
  stat_smooth(method = "loess", span=10, se=FALSE) +
  scale_y_log10()+
scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  labs(x = "Nodule Surface Area(cm2)", y = "CFU/Nodule")+
    facet_wrap(~ Host, nrow = 3, labeller = labeller(Host = host_labels))+
  plot_theme()+
  theme(legend.position = "none")

#dev.off()


#Grid format of Figure 2
Figure_2_plot<- (surface_area_plot / rhzreleasedbr_plot / rhzreleased_plot)+
  plot_layout(guides="collect", heights=c(1,1,1.75)) + 
  theme(legend.position = "right")+
  plot_annotation(tag_levels = "A") &
  theme(plot.margin= unit(c(0.5, 0.5, 0.5, 0.5), "pt"),
        panel.spacing    = unit(1.5, "mm"),
        legend.box.margin = margin(0, 0, 0, 0, unit = "pt"),
        plot.tag = element_text(size = 26, face = "bold", hjust = 6, vjust= -2))


svg(filename=file.path(plot_dir,"Figure_2_ver6.svg"),width=10, height=14)
print(Figure_2_plot)
dev.off()


```

##Figure 3
```{r Figure 3, echo=FALSE, warning=FALSE, message=TRUE}
#Plots generated for figure 3. Single plot exporting has been "#" out to make this knit, if an export is needed removing the "#" from the svg() and dev.off() functions below.

#Mean Host Biomass Plot
#svg(filename=file.path(plot_dir,"Figure_3_biomass.svg"), width=3, height=3)
biomass_plot<-
ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_biomass_g, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(x="", y="Mean Host Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Nodule Count Plot
#svg(filename=file.path(plot_dir,"Figure_3_nodulecount.svg"), width=3, height=3)
nodulecount_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_nod_num, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(y="Mean Nodule Count", x="Mean Proportion of Branched Nodules",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Prop Nodules across Host Variety
#svg(filename=file.path(plot_dir,"Figure_3_propbranch.svg"), width=5, height=5)
proplobe_plot<-ggplot(str.host_mean_df, aes(x=Variety, y=mean_prop_lobe, group=Strain, color=Variety))+
  geom_line(color="gray",linewidth=1, alpha=0.3, show.legend = FALSE)+
  geom_jitter(alpha=0.7, width=0.25, size=3, show.legend = FALSE)+
  labs(x="Alfalfa Host Variety", y="Mean Proportion of\nBranched Nodules")+
    plot_theme()
#dev.off()

# Generates the combined figure 3 plot
fig_3_plot <- biomass_plot + nodulecount_plot + proplobe_plot +
  plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") &
  theme(legend.position = "top", plot.tag = element_text(size = 26, face = "bold", hjust = 0.2, vjust = -4)) 


svg(filename=file.path(plot_dir,"Figure_3.svg"), width=15, height=7)
print(fig_3_plot)
dev.off()
```

#Supplemental

```{r Supplement tables}
#Building supplemental table of anova results from tested models
tab.1a <- data.frame(dataset="cfu_df", model="log10(Nodule.area)~Nodule.lobes*Host*Strain+rep",Term=row.names(aov.nodarea),
                     Df=aov.nodarea$Df,
                     Prop.Var=round(aov.nodarea$`Sum Sq`/sum(aov.nodarea$`Sum Sq`),3),
                     Fstat=round(aov.nodarea$`F value`,2),
                     Pvalue=aov.nodarea$`Pr(>F)`)
tab.1b <- data.frame(dataset="cfu_df", model="log10(rhz_released)~as.factor(Nodule.lobes)*Host+as.factor(Strain)+rep",Term=row.names(aov.rhzreleased),
                     Df=aov.rhzreleased$Df,
                     Prop.Var=round(aov.rhzreleased$`Sum Sq`/sum(aov.rhzreleased$`Sum Sq`),3),
                     Fstat=round(aov.rhzreleased$`F value`,2),
                     Pvalue=aov.rhzreleased$`Pr(>F)`)
tab.2a <- data.frame(dataset="str.host_mean_df", model="mean_nod_num~mean_prop_lobe*Variety+Strain",Term=row.names(aov.nodnum),
                     Df=aov.nodnum$Df,
                     Prop.Var=round(aov.nodnum$`Sum Sq`/sum(aov.nodnum$`Sum Sq`),3),
                     Fstat=round(aov.nodnum$`F value`,2),
                     Pvalue=aov.nodnum$`Pr(>F)`)
tab.2c <- data.frame(dataset="str.host_mean_df", model="mean_biomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.biomass),
                     Df=aov.biomass$Df,
                     Prop.Var=round(aov.biomass$`Sum Sq`/sum(aov.biomass$`Sum Sq`),3),
                     Fstat=round(aov.biomass$`F value`,2),
                     Pvalue=aov.biomass$`Pr(>F)`)
tab.s1 <- data.frame(dataset="rhztestarea", model="log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep",Term=row.names(aov.rhztestarea),
                     Df=aov.rhztestarea$Df,
                     Prop.Var=round(aov.rhztestarea$`Sum Sq`/sum(aov.rhztestarea$`Sum Sq`),3),
                     Fstat=round(aov.rhztestarea$`F value`,2),
                     Pvalue=aov.rhztestarea$`Pr(>F)`)
supplement_table <- rbind(tab.1a, tab.1b, tab.2a, tab.2c, tab.s1)
write.csv(supplement_table, file=paste0(table_dir, "Supplement_table1.csv"))

```



## Mariale trying to figure out how to better represent the span of shoot biomass 

```{r}

### Phylogeny

## These will use Sohini's paths 

library(tidyverse)
library(ggpubr)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(ggtree)
library(vegan)
library(corrplot)
library(patchwork)
#knitr::opts_chunk$set(message = FALSE)
#install.packages("ggnewscale")
library(ggnewscale)
library(ape)
library(readxl)
#install.packages("BiocManager")
#BiocManager::install(c("ggtree", "ggtreeExtra"))
library(ggtreeExtra)
#install.packages("colorRamps")
library(colorRamps)


ss_df<-data.frame(fread("../data/MasterSpreadsheet_AVT.csv"))
#View(ss_df)
# Removing erroneous columns in the original data frame, this includes collection notes
ss_df <- ss_df[, !colnames(ss_df) %in% "NOTES"]
#View(ss_df)

# Adding total nodule count and the proportion of branched nodules
ss_df<-ss_df %>%
  mutate(Nod_count=Nod_lobe+Nod_unlobed)%>% #Counting total observed nodules
  mutate(prop_lobe=Nod_lobe/Nod_count) %>% #Creating the proportion of branched nodules to nodule count
  mutate(Biomass_g=Rootweight_g+Shootweight_g) #Coutning total host mass biomass
## I am going to filter out MIX

ss_df <- ss_df %>% 
  filter(!Strain %in% c("MIX", "NIG", "MC2", "EXC", "UNI")) %>% #Maria: Liz here you can see an example of how I get rid of the treatments that are not useful for this analysis.
    filter(!is.na(Strain), Strain != "")


#View(ss_df)
## I am only going to select the things I am interested in 

ss_df_shoot <- ss_df %>% select(Strain, Variety, Shootweight_g)
#View(ss_df)


#ss_df_shoot <- ss_df %>% select(Strain, Variety, Shootweight_g) ## According to her code the mean of shoot biomass is named mean_abiomass_g
ss_df_shoot_mean <- ss_df_shoot %>%
  group_by(Strain, Variety) %>%
  summarise(shoot_ave = mean(Shootweight_g, na.rm = TRUE))

View(ss_df_shoot_mean)

Shoot_biomass <- ss_df_shoot_mean %>%
  pivot_wider(
    names_from = Variety,   # Variety values become column names
    values_from = shoot_ave   # Biomass values fill the cells
  )

Shoot_biomass

```

## For the actual tree

The aesthetics of the tree were mostly arranged in the itols app. However, I created the metadata to be imported to itols in the code bellow
```{r}


### These files were done by Sohini. 

tree<-read.tree("/Users/mariaalejandragilpolo/OneDrive - The Pennsylvania State University/Burghardt, Liana T's files - GitHub/AVT_SingleInoculation/Data/AVT_single_copy_core_concat.fasta.treefile")
tree

# 2. Read your annotation file from iTOL
# Replace the path with wherever you saved Annotation_iTOL_final.txt
meta_raw <- read.delim("/Users/mariaalejandragilpolo/OneDrive - The Pennsylvania State University/Burghardt, Liana T's files - GitHub/AVT_SingleInoculation/Data/Annotation_iTOL_final.txt", sep = "\t", header = TRUE)
meta_raw ## This is the meta_raw used in Sohini's paper: Morphology and origin

### Me and Sohini have different labels. She uses AVT1.. and I use 001... I am trying to work with the codes that I used 

Codes <- read_excel("/Users/mariaalejandragilpolo/Downloads/Strain_codes_ForTree.xlsx") #Has all the 117 strains


## Rename Strain as Strain_Code

Shoot_biomass_edit <- Shoot_biomass %>% dplyr::rename(Strain_Code = Strain) 

Shoot_biomass_withCodes <- left_join(Codes, Shoot_biomass_edit, by = "Strain_Code") 


### Start with the labels 


#### labels_df for labels. 
labels_df <- data.frame(
  tip = meta_tree$Strain,
  new_label = meta_tree$Strain_Code
)
labels_df <- labels_df[!is.na(labels_df$tip) & !is.na(labels_df$new_label), ]
labels_df


### Create a dataframe with the datapoints I care about: The aboveground biomass data is in Shoot_biomass

meta_tree <- meta_raw %>%
  left_join(
    Shoot_biomass_withCodes %>%
      select(Strain, Strain_Code, O, V, S), 
    by = "Strain"
  ) %>%
  select(-Morphology, -Source_variety)

#View(meta_tree) Looking good

### Now I will write them in a format for itols 

cat("LABELS\nSEPARATOR TAB\n\nDATA\n",
    file = "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/nodule_branching_morphology/Tables/labels.txt")
write.table(labels_df, "/Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/nodule_branching_morphology/Tables/labels.txt",
            sep = "\t", append = TRUE, row.names = FALSE,
            col.names = FALSE, quote = FALSE)


# Generate iTOL DATASET_SYMBOL file: Shape plot
# The idea is to have three colored circles per strain, one per host (O, V, S) that will have different sizes depeding on how big the raw biomass is.
# The circle size is raw biomass on that host.
# Circle color = host (O = red, V = pinkish, S = blue).## I am using the same colors that Liz used
# Input:  meta_tree with columns Strain, O, V, S
# Output: itol_annotations/biomass_symbols.txt

host_colors <- c(
  "O" = "#471323",   
  "V" = "#FCB0B3",  
  "S" = "#ADB9E3"   
)
 
# Create the file

 
dir.create("itol_annotations", showWarnings = FALSE)
 
df <- meta_tree   # I am using df to replace meta_tree 
 

## Build the table that will be the input in itols. 
# DATASET_EXTERNALSHAPE expects: ID  value_field1  value_field2  value_field3
# Sizes are scaled automatically across all values in the dataset.
 
shape_df <- df[, c("Strain", "O", "V", "S")]
shape_df
shape_df <- shape_df[!is.na(shape_df$Strain) & nzchar(shape_df$Strain), ]
 shape_df

 outfile <- "itol_annotations/biomass_externalshape.txt"

 ## Writing the file. Dataset_externalshape is important to keep them separate 
cat(
  "DATASET_EXTERNALSHAPE\n",
  "SEPARATOR TAB\n",
  "DATASET_LABEL\tShoot biomass\n",
  "COLOR\t#000000\n",
 
  # One column per host, with column headers
  "FIELD_LABELS\tO\tV\tS\n",
  "FIELD_COLORS\t", host_colors["O"], "\t",
                    host_colors["V"], "\t",
                    host_colors["S"], "\n",
 
  # Shape per column: 2 = filled circle (per iTOL docs)
  "FIELD_SHAPES\t2\t2\t2\n",
 
  "SHAPE_SCALING\t1\n",        # scale shapes by value (the whole point)
  "SHOW_VALUES\t0\n",          
 
  # Visual spacing
  "SHAPE_SPACING\t10\n",       # gap between columns
  "HORIZONTAL_GRID\t1\n",
  "VERTICAL_GRID\t1\n",
 
  # Legend
  "LEGEND_TITLE\tBiomass by host\n",
  "LEGEND_SHAPES\t2\t2\t2\n",
  "LEGEND_COLORS\t", host_colors["O"], "\t",
                     host_colors["V"], "\t",
                     host_colors["S"], "\n",
  "LEGEND_LABELS\tO\tV\tS\n",
 
  "\nDATA\n",
  sep = "",
  file = outfile
)
 
write.table(shape_df, outfile,
            sep = "\t", append = TRUE,
            row.names = FALSE, col.names = FALSE,
            quote = FALSE, na = "X")
 
cat("Wrote:", outfile, "\n")
cat("Drag it onto your tree in iTOL.\n")
cat("If circles are too big/small, adjust via the gear icon next to the dataset.\n")
 


```
