---
title: "Nodule branching, size, and symbiosis outcomes are shaped by natural genetic variation in rhizobial bacteria and alfalfa hosts "
author: "Elizabeth Leslie Paillan - elpailla@ncsu.edu & Liana Burghardt - liana.burghardt@psu.edu"
date: "2025-10-25"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

Nodule traits being examined: Proportion Lobed (prop_lobe= \# of lobed nodule/ total \# nodule sample) a lower proportion means fewer lobed nodules, and a higher proportion means larger number of nodules. Nodule Prevalence (nod_prev=total \# nodule sample), the term origniates from the avt_field experiement where I was dividing the number of nodules found by plant, that is not needed since its was one plant (mostly), for this experiment its just the total nodule number.

#Libraries Used

```{r Library setup, echo = TRUE, warning=FALSE}
library(tidyverse); packageVersion("tidyverse")
library(BSDA); packageVersion("BSDA")
library(patchwork); packageVersion("patchwork")
library(svglite); packageVersion("svglite")
library(performance); packageVersion("performance")
library(data.table); packageVersion("data.table")
set.seed(100)
```

```{r GGplot theme, include=FALSE}
#Below is the general theme used for all ggplot2 figures
plot_theme <- function(base_size = 18) {
  theme_bw(base_size = base_size) +
    theme(legend.title = element_text(face = "bold"),
      legend.text  = element_text(face = "bold"),
      axis.title   = element_text(face = "bold"),
      strip.text   = element_text(face = "bold"),
      panel.grid   = element_blank(),
      panel.background = element_blank(),
      axis.line    = element_line(linewidth = 1.05, color = "black"))
}

#Changing host variety names to their full names & add designated color scheme
host_labels <- c(O = "Onedia", S = "SW402", V = "Vernal")
host_colors <- c(O = "#471323", S = "#ADB9E3", V = "#FCB0B3")

#same thing while nodule branch number
nbranch_labels <- c("1", "2", "3-4", "5+")
nbranch_colors <- c("1" = "#273E47",
                    "2" = "#BAA5FF",
                    "3-4" = "#6E9887",
                    "5+" = "#A5402D")
nbranch_shapes <- c("1" = 16, "2" = 15, "3-4" = 17, "5+" = 18)

#and strain ID
strain_shapes <- modifyList(as.list(setNames(rep(16, length(unique(str.host_mean_df$Strain))),
                   unique(str.host_mean_df$Strain))),
                   list("001"=15, "012"=12, "044"=17, "059"=18, "070"=8, "077"=4, "112"=25, "117"=3))

strain_labels <- c("001"=1, "012"=2, "044"=3, "059"=4, "070"=5, "077"=6, "112"=7, "117"=8)
strain_shapes <- c("001"=15, "012"=24, "044"=17, "059"=18, "070"=8, "077"=4, "112"=25, "117"=3)



```

#Functions written for the analysis:

```{r Function: Anova on linear models -> pulling the summary and r2, echo=TRUE}
### This function runs a general anova, and spits out the summary() table and export it
lm_run<- function(model, model_name, save_path=getwd()) {
  anova_output<- anova(model)
  model_summary <- summary(model)
  results_list<- list()
  
  results_list[[paste0(model_name, "_anova")]] <- anova_output
  results_list[[paste0(model_name, "_summary")]] <- model_summary
  
  anova_df<- as.data.frame(anova_output)
  summary_df<- as.data.frame(model_summary$coefficients)
  
  return(results_list)
}
```

```{r Function: Detecting and Listing Outliers, }
#This function is used to detect outliers, create a list of the identified outliers.
detectview_outlier <- function(data, variable) {
  outliers <- boxplot(data[[variable]], plot = FALSE)$out #Generates a barplot to ID outliers
  outlier_data <- data[data[[variable]] %in% outliers, ]
  
  outlier_plot <- ggplot(data, aes(x = 1, y = !!rlang::sym(variable))) + 
    geom_boxplot() + 
    geom_point(data = outlier_data, aes(x = 1, y = !!rlang::sym(variable)), color = "red") + #Makes the outlier more visible
    labs(x = "", y = variable, title = paste("Boxplot of", variable, "outliers"))
  
  return(list(plot = outlier_plot, outlier_data = outlier_data)) #Provides detail about the identified outliers
}
```

#Importing Data

```{r Data Import of Single Inoculation (ss_df), results = HIDE, warning = HIDE, collapse=TRUE}
#Import and tidying of Data
#setwd("~/GitHub/nodule_branching_morphology/Analysis/")

#Setting up plot and directories to be easily found:
plot_dir<-("../Figures/")
table_dir<-("../Tables/")

ss_df<-data.frame(fread("../data/MasterSpreadsheet_AVT.csv"))

# Removing erroneous columns in the original data frame, this includes collection notes
ss_df <- ss_df[, !colnames(ss_df) %in% "NOTES"]

# Adding total nodule count and the proportion of branched nodules
ss_df<-ss_df %>%
  dplyr::select(Strain, Variety, Source_variety, Rep, Rootweight_g, Shootweight_g, Nod_lobe, Nod_unlobed)%>% #Selecting columns relevant to the analysis
  mutate(Nod_count=Nod_lobe+Nod_unlobed)%>% #Counting total observed nodules
  mutate(prop_lobe=Nod_lobe/Nod_count) %>% #Creating the proportion of branched nodules to nodule count
  filter_at(vars(Nod_count:prop_lobe), all_vars(!is.infinite(.)))%>% #Checking for any infinite values
  mutate(Biomass_g=Rootweight_g+Shootweight_g)%>% #Counting total host mass biomass
  dplyr::filter(!Strain %in% c("MIX", "NIG", "MC2", "EXC", "UNI"))%>%
  drop_na() #Removing any NAs


# Using the detect outlier function to isolate them for view if needed 
Biom_outlier<-detectview_outlier(ss_df,"Biomass_g")
Lobe_outlier<-detectview_outlier(ss_df,"prop_lobe")

#Prop_lobe.outliers<-c("110_O_a")
#ss_df<-ss_df %>%
  dplyr::filter(!(Label %in% Prop_lobe.outliers))

# Separate data frame that is the mean of every Isolate-Host variety combination across all replicates
Mean_lobe<-ss_df %>% 
            group_by(Strain,Variety,Source_variety)%>% 
            dplyr::summarize(lobe_mean=mean(prop_lobe, na.rm=TRUE))
```

```{r Data Import of CFU experiment (cfu_df), message=FALSE, warning=FALSE, page.print=FALSE}
#Loading in the cfu dataset
cfu_df<-data.frame(fread("../data/AVT_CFU_count.csv"))

#Removing erroneous columns from the data frame like notes from the harvest, and renames one column to be easier to call into R
cfu_df <- cfu_df %>%
  mutate(Comments=NULL)%>%
  mutate(dilution.counted=NULL)%>%
  rename(rhz_released = Lianas.rhizobia.released.per.nodule.calculation)%>% 
#Since we had a couple of nodules that had higher than 5 branches we combined them into a 5+ category
  mutate(Nodule.lobes = ifelse(Nodule.lobes == "more than 5", "5+", 
                        ifelse(Nodule.lobes == "More than 5", "5+",
                        ifelse(Nodule.lobes == "5", "5+",
#Few nodules had 3-4 branches so we combined them into a 3-4 category
                        ifelse(Nodule.lobes %in% c("3", "4"), "3-4", Nodule.lobes)))))%>%
#Making sure the branch coulmn is considered a factor
  mutate(Nodule.lobes = as.factor(Nodule.lobes))

#Extracting sample data from the sample-ID
cfu_df<-separate(cfu_df, Plant, into = c("Strain","Host","rep"), sep="\\.")

#Ensuring intended variables are numeric and factor as intended
cfu_df$rhz_released<-as.numeric(cfu_df$rhz_released)
cfu_df$Nodule.area<-as.numeric(cfu_df$Nodule.area)
cfu_df$spot.avg <- as.numeric(gsub("[^0-9.]", "", cfu_df$spot.avg))


#Detecting outliers in the CFU data for view if needed (not used)
RhzReleased_outlier<-detectview_outlier(cfu_df,"rhz_released")
Nodule.area_outlier<-detectview_outlier(cfu_df,"Nodule.area")

#Removing 0.44.V.b.4  because of a sampling error making it a massive outlier in the nodule.area measure
Nodule.area.outliers<-c("044.V.b.4")
RhzReleased.outliers<-c("070.O.b.5")

cfu_df<-cfu_df %>%
#Removing size that was likely an error during processing
    filter(!(Ind.Label %in% RhzReleased.outliers))%>%
  filter(!(Ind.Label %in% Nodule.area.outliers))

#Removing any NA's introduced by adding a log transformation to the colony counts
cfu_df[is.infinite(log10(cfu_df$rhz_released)) | 
         is.na(log10(cfu_df$rhz_released)), ]
cfu_df <- subset(cfu_df, rhz_released > 0 & !is.na(rhz_released))

```

```{r Strain Reaction norm df (str.host_mean_df), collapse=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Another dataframe to create a reaction norm like plot
#Taking the mean across reps, strain, and variety to reduce noise
str.host_mean_df <- ss_df %>%
  group_by(Strain,Variety)%>%
  summarise(mean_prop_lobe=mean(prop_lobe,na.rm=TRUE), 
            mean_abvmass_g=mean(Shootweight_g,na.rm=TRUE),
            mean_belmass_g=mean(Rootweight_g, na.rm=TRUE),
            mean_biomass_g=mean(Biomass_g, na.rm=TRUE),
            mean_nod_num=mean(Nod_count, na.rm=TRUE),
            mean_agbiomass_g=mean(Shootweight_g, na.rm=TRUE))
```

#Statistics used 
##Rhizboial focused traits 
```{r Rhizobial Trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova examining Nodule area response predicted by branch #, host variety,and isolate ID
nodarea<-lm(log10(Nodule.area)~Nodule.lobes*Host*Strain+rep, data=cfu_df)
nodarea.results<-lm_run(nodarea, model_name="nodarea")
aov.nodarea<-anova(nodarea)
write.csv(aov.nodarea, file=paste(table_dir,"Nodule-Area-response_anova.csv"))
aov.nodarea
aov.nodarea[["Sum Sq"]] / sum(aov.nodarea[["Sum Sq"]])

#Model examining Amount of viable rhizobia released focused on number of nodule branches as a predictor
rhz_released<-lm(log10(rhz_released)~Nodule.lobes*Host+as.factor(Strain)+rep, data=cfu_df)
rhz_released.results<-lm_run(rhz_released, "rhzreleased")
aov.rhzreleased<- anova(rhz_released)
write.csv(aov.rhzreleased, file=paste(table_dir,"RhizobiaReleased-response_anova.csv"))
aov.rhzreleased
aov.rhzreleased[["Sum Sq"]] / sum(aov.rhzreleased[["Sum Sq"]])

# Model examining amount of rhizobia released focused on nodule surface area as a predictor 
rhztestarea<-lm(log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep, data=cfu_df)
rhztestarea.results<-lm_run(rhztestarea,"rhzarea")
aov.rhztestarea<- anova(rhztestarea)
write.csv(aov.rhztestarea, file=paste(table_dir,"Nodulearea-by-surfacearea_anova.csv"))
aov.rhztestarea
aov.rhztestarea[["Sum Sq"]] / sum(aov.rhztestarea[["Sum Sq"]])
```

##Statistical Analysis of the Host plant traits
```{r host plant trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova is being ran on models associated with host traits 
#Modeling the mean nodule number predicted by proportion of branched nodules and host variety
nodnumber<-lm(mean_nod_num~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
nodnum.results<-lm_run(nodnumber, "nodnum")
aov.nodnum<-anova(nodnumber)
write.csv(aov.nodnum, file=paste(table_dir,"Nodulenumber-response_anova.csv"))
aov.nodnum
aov.nodnum[["Sum Sq"]] / sum(aov.nodnum[["Sum Sq"]])

#Modeling the mean host plant biomass predicted by the mean proportion of branched nodules and host variety
biomass<-lm(mean_biomass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
biomass.results<-lm_run(biomass, "biomass")
aov.biomass<-anova(biomass)
write.csv(aov.biomass, file=paste(table_dir,"Biomass-response_anova.csv"))
aov.biomass
aov.biomass[["Sum Sq"]] / sum(aov.biomass[["Sum Sq"]])

##proportion branches
propbr<- lm(prop_lobe ~ Variety*Strain+Rep,data=ss_df)
propbr.results<- lm_run(propbr,"meanbranch")
aov.propbr<-anova(propbr)
write.csv(aov.propbr, file=paste(table_dir,"proplobe-response_anova.csv"))
aov.propbr
aov.propbr[["Sum Sq"]] / sum(aov.propbr[["Sum Sq"]])


#Modeling the mean host above and below ground biomass separately
##above ground
a.biomass<-lm(mean_abvmass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
a.biomass.results<-lm_run(a.biomass, "biomass")
aov.a.biomass<-anova(a.biomass)
write.csv(aov.a.biomass, file=paste(table_dir,"Abovebround-biomass-response_anova.csv"))
aov.a.biomass
aov.a.biomass[["Sum Sq"]] / sum(aov.a.biomass[["Sum Sq"]])

##below ground
b.biomass<-lm(mean_belmass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
b.biomass.results<-lm_run(b.biomass, "Belowground biomass")
aov.b.biomass<-anova(b.biomass)
write.csv(aov.b.biomass, file=paste(table_dir,"Belowground-biomass-response_anova.csv"))
aov.b.biomass
aov.b.biomass[["Sum Sq"]] / sum(aov.b.biomass[["Sum Sq"]])
```

#Figures
##Figure 2
```{r Figure 2, echo=FALSE, warning=FALSE, message=TRUE}
#Single plot exporting has been "#" out to make this knit, if an export is needed removing the "#" from the pdf() and dev.off() below.

pdf(file=file.path(plot_dir,"Figure_2_surfacearea.pdf"), width=3, height=3)
surface_area_plot<- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=Nodule.area, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black")+
  labs(x="", y="Nodule Surface Area(cm2)", shape="Strain", color="Number of\nNodule\nBranches")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  plot_theme()
dev.off()

pdf(file=file.path(plot_dir,"Figure_2_released_rhizobia_branch.pdf"), width=3, height=3)
rhzreleasedbr_plot <- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=rhz_released, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black")+
  scale_y_log10()+
  labs(x= "Number of Nodule Branches", y="CFU/Nodule", shape="Strain", color="Host Variety")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  plot_theme()+
  theme(legend.position = "none")
  
dev.off()

pdf(file=file.path(plot_dir,"Figure_2_released_rhizobia_branch_ver5.pdf"), width=3, height=3)
rhzreleased_plot<-
  ggplot(cfu_df, aes(x = Nodule.area, y = rhz_released, color = Nodule.lobes)) +
  geom_point(size=3)+
  stat_smooth(method = "loess", span=10, se=FALSE) +
  scale_y_log10()+
scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  labs(x = "Nodule Surface Area(cm2)", y = "CFU/Nodule")+
    facet_wrap(~ Host, nrow = 3, labeller = labeller(Host = host_labels))+
  plot_theme()+
  theme(legend.position = "none")

dev.off()


#Grid format of Figure 2
Figure_2_plot<- (surface_area_plot / rhzreleasedbr_plot / rhzreleased_plot)+
  plot_layout(guides="collect", heights=c(1,1,1.75)) + 
  theme(legend.position = "right")+
  plot_annotation(tag_levels = "A") &
  theme(plot.margin= unit(c(0.5, 0.5, 0.5, 0.5), "pt"),
        panel.spacing    = unit(1.5, "mm"),
        legend.box.margin = margin(0, 0, 0, 0, unit = "pt"),
        plot.tag = element_text(size = 26, face = "bold", hjust = 6, vjust= -2))


svg(filename=file.path(plot_dir,"Figure_2.svg"), width=13, height=13)
print(Figure_2_plot)
dev.off()

pdf(file=file.path(plot_dir,"Figure_2_ver6.pdf"),width=10, height=14)
print(Figure_2_plot)
dev.off()
```

##Figure 3
```{r Figure 3, echo=FALSE, warning=FALSE, message=TRUE}
#Plots generated for figure 3. Single plot exporting has been "#" out to make this knit, if an export is needed remove the "#" from the pdf() and dev.off() functions below.

#Mean Host Biomass Plot
#pdf(filename=file.path(plot_dir,"Figure_3_biomass.pdf"), width=3, height=3)
biomass_plot<-
ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_biomass_g, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(x="", y="Mean Host Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Host Shoot Biomass plot
#pdf(filename=file.path(plot_dir,"Figure_3_shoot_biomass.pdf"), width=3, height=3)
abiomass_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_abvmass_g, color=Variety))+
  geom_point(aes(shape=Strain),alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(x="", y="Mean Host\nShoot Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

#Mean Host Root Biomass plot
#pdf(filename=file.path(plot_dir,"Figure_3_root_biomass.pdf"), width=3, height=3)
b.biomass_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_belmass_g, color=Variety))+
  geom_point(aes(shape=Strain) ,alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(x="Mean Proportion of Branched Nodules ", y="Mean Host\nRoot Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

#Mean Nodule Count Plot
#pdf(filename=file.path(plot_dir,"Figure_3_nodulecount.pdf"), width=3, height=3)
nodulecount_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_nod_num, color=Variety))+
  geom_point(aes(shape=Strain), alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(y="Mean Nodule Count", x="Mean Proportion of Branched Nodules",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Prop Nodules across Host Variety
#pdf(filename=file.path(plot_dir,"Figure_3_propbranch.pdf"), width=5, height=5)
proplobe_plot<-
  ggplot(str.host_mean_df, aes(x=Variety, y=mean_prop_lobe, group=Strain, color=Variety))+
  geom_line(color="gray",linewidth=1, alpha=0.3, show.legend = FALSE)+
  geom_jitter(aes(shape=Strain),alpha=0.7, width=0.25, size=3, show.legend = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(x="Alfalfa Host Variety", y="Mean Proportion of\nBranched Nodules")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

# Generates the combined figure 3 plot
fi3panelA<- abiomass_plot + b.biomass_plot + plot_layout(ncol=2)

fig_3_plot <- fi3panelA  + nodulecount_plot + proplobe_plot +
  plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", plot.tag = element_text(size = 26, face = "bold", hjust = 0.2, vjust = -4)) 


svg(filename=file.path(plot_dir,"Figure_3.svg"), width=13, height=13)
print(fig_3_plot)
dev.off()

#Doesnt have the fixed axes label and "other" label in legend
pdf(file=file.path(plot_dir,"Figure_3.pdf"), width=10, height=10)
print(fig_3_plot)
dev.off()
```

#Supplemental

```{r Supplement tables}
#Building supplemental table of anova results from tested models
tab.1a <- data.frame(dataset="cfu_df", model="log10(Nodule.area)~Nodule.lobes*Host*Strain+rep",Term=row.names(aov.nodarea),
                     Df=aov.nodarea$Df,
                     Prop.Var=round(aov.nodarea$`Sum Sq`/sum(aov.nodarea$`Sum Sq`),3),
                     Fstat=round(aov.nodarea$`F value`,2),
                     Pvalue=aov.nodarea$`Pr(>F)`)
tab.1b <- data.frame(dataset="cfu_df", model="log10(rhz_released)~as.factor(Nodule.lobes)*Host+as.factor(Strain)+rep",Term=row.names(aov.rhzreleased),
                     Df=aov.rhzreleased$Df,
                     Prop.Var=round(aov.rhzreleased$`Sum Sq`/sum(aov.rhzreleased$`Sum Sq`),3),
                     Fstat=round(aov.rhzreleased$`F value`,2),
                     Pvalue=aov.rhzreleased$`Pr(>F)`)
tab.2a <- data.frame(dataset="str.host_mean_df", model="mean_nod_num~mean_prop_lobe*Variety+Strain",Term=row.names(aov.nodnum),
                     Df=aov.nodnum$Df,
                     Prop.Var=round(aov.nodnum$`Sum Sq`/sum(aov.nodnum$`Sum Sq`),3),
                     Fstat=round(aov.nodnum$`F value`,2),
                     Pvalue=aov.nodnum$`Pr(>F)`)
tab.2c <- data.frame(dataset="str.host_mean_df", model="mean_abovebiomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.a.biomass),
                     Df=aov.a.biomass$Df,
                     Prop.Var=round(aov.a.biomass$`Sum Sq`/sum(aov.a.biomass$`Sum Sq`),3),
                     Fstat=round(aov.a.biomass$`F value`,2),
                     Pvalue=aov.a.biomass$`Pr(>F)`)
tab.2z <- data.frame(dataset="str.host_mean_df", model="mean_belowbiomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.b.biomass),
                     Df=aov.b.biomass$Df,
                     Prop.Var=round(aov.b.biomass$`Sum Sq`/sum(aov.b.biomass$`Sum Sq`),3),
                     Fstat=round(aov.b.biomass$`F value`,2),
                     Pvalue=aov.b.biomass$`Pr(>F)`)
tab.s1 <- data.frame(dataset="rhztestarea", model="log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep",Term=row.names(aov.rhztestarea),
                     Df=aov.rhztestarea$Df,
                     Prop.Var=round(aov.rhztestarea$`Sum Sq`/sum(aov.rhztestarea$`Sum Sq`),3),
                     Fstat=round(aov.rhztestarea$`F value`,2),
                     Pvalue=aov.rhztestarea$`Pr(>F)`)
supplement_table <- rbind(tab.1a, tab.1b, tab.2a, tab.2c,tab.2z, tab.s1)
write.csv(supplement_table, file=paste0(table_dir, "Supplement_table1.csv"))

```
