---
title: "Nodule branching, size, and symbiosis outcomes are shaped by natural genetic variation in rhizobial bacteria and alfalfa hosts "
author: "Elizabeth Leslie Paillan - elpailla@ncsu.edu & Liana Burghardt - liana.burghardt@psu.edu"
date: "2025-10-25"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

Nodule traits being examined: Proportion Lobed (prop_lobe= \# of lobed nodule/ total \# nodule sample) a lower proportion means fewer lobed nodules, and a higher proportion means larger number of nodules. Nodule Prevalence (nod_prev=total \# nodule sample), the term origniates from the avt_field experiement where I was dividing the number of nodules found by plant, that is not needed since its was one plant (mostly), for this experiment its just the total nodule number.

#Libraries Used

```{r Library setup, echo = TRUE, warning=FALSE}
library(tidyverse); packageVersion("tidyverse")
library(BSDA); packageVersion("BSDA")
library(patchwork); packageVersion("patchwork")
library(svglite); packageVersion("svglite")
library(performance); packageVersion("performance")
library(data.table); packageVersion("data.table")
library(here); packageVersion("here")
set.seed(100)
```

```{r GGplot theme, include=FALSE}
#Below is the general theme used for all ggplot2 figures
plot_theme <- function(base_size = 18) {
  theme_bw(base_size = base_size) +
    theme(legend.title = element_text(face = "bold"),
      legend.text  = element_text(face = "bold"),
      axis.title   = element_text(face = "bold"),
      strip.text   = element_text(face = "bold"),
      panel.grid   = element_blank(),
      panel.background = element_blank(),
      axis.line    = element_line(linewidth = 1.05, color = "black"))
}

#Changing host variety names to their full names & add designated color scheme
host_labels <- c(O = "Onedia", S = "SW402", V = "Vernal")
host_colors <- c(O = "#471323", S = "#ADB9E3", V = "#FCB0B3")

#same thing while nodule branch number
nbranch_labels <- c("1", "2", "3-4", "5+")
nbranch_colors <- c("1" = "#273E47",
                    "2" = "#BAA5FF",
                    "3-4" = "#6E9887",
                    "5+" = "#A5402D")
nbranch_shapes <- c("1" = 16, "2" = 15, "3-4" = 17, "5+" = 18)

#and strain ID

strain_labels <- c("001"=1, "012"=2, "044"=3, "059"=4, "070"=5, "077"=6, "112"=7, "117"=8)
strain_shapes <- c("001"=15, "012"=24, "044"=17, "059"=18, "070"=8, "077"=4, "112"=25, "117"=3)



```

#Functions written for the analysis:

```{r Function: Anova on linear models -> pulling the summary and r2, echo=TRUE}
### This function runs a general anova, and spits out the summary() table and export it
lm_run<- function(model, model_name, save_path=getwd()) {
  anova_output<- anova(model)
  model_summary <- summary(model)
  results_list<- list()
  
  results_list[[paste0(model_name, "_anova")]] <- anova_output
  results_list[[paste0(model_name, "_summary")]] <- model_summary
  
  anova_df<- as.data.frame(anova_output)
  summary_df<- as.data.frame(model_summary$coefficients)
  
  return(results_list)
}
```

#Importing Data

```{r Data Import of Single Inoculation (ss_df), results = HIDE, warning = HIDE, collapse=TRUE}
#Import and tidying of Data
here::i_am("nodule_branching_morphology/Analysis/Nodule_branching.Rmd")

#Setting up plot and directories to be easily found:
plot_dir<- here("nodule_branching_morphology", "Figures") 
table_dir<-here("nodule_branching_morphology", "Tables")
ss_df<-data.frame(fread(here("nodule_branching_morphology", "Data","MasterSpreadsheet_AVT.csv")))

# Removing erroneous columns in the original data frame, this includes collection notes
ss_df <- ss_df[, !colnames(ss_df) %in% "NOTES"]

# Adding total nodule count and the proportion of branched nodules
ss_df<-ss_df %>%
  dplyr::select(Strain, Variety, Source_variety, Rep, Rootweight_g, Shootweight_g, Nod_lobe, Nod_unlobed, Label)%>% #Selecting columns relevant to the analysis
  mutate(Nod_count=Nod_lobe+Nod_unlobed)%>% #Counting total observed nodules
  mutate(prop_lobe=Nod_lobe/Nod_count) %>% #Creating the proportion of branched nodules to nodule count
  filter_at(vars(Nod_count:prop_lobe), all_vars(!is.infinite(.)))%>% #Checking for any infinite values
  mutate(Biomass_g=Rootweight_g+Shootweight_g)%>% #Counting total host mass biomass
  dplyr::filter(!Strain %in% c("MIX", "NIG", "MC2", "EXC", "UNI"))%>%
  drop_na()

# Removing sampling error
prop_lobe.outliers<-c("110_O_a")
ss_df<-ss_df %>%
  dplyr::filter(!(Label %in% prop_lobe.outliers))

```

```{r Data Import of CFU experiment (cfu_df), message=FALSE, warning=FALSE, page.print=FALSE}
#Loading in the cfu dataset
cfu_df<-data.frame(fread(here("nodule_branching_morphology","Data","AVT_CFU_count.csv")))

cfu_df <- cfu_df %>% 
#Since we had a couple of nodules that had higher than 5 branches we combined them into a 5+ category
  mutate(Nodule.lobes = ifelse(Nodule.lobes == "more than 5", "5+", 
                        ifelse(Nodule.lobes == "More than 5", "5+",
                        ifelse(Nodule.lobes == "5", "5+",
#Few nodules had 3-4 branches so we combined them into a 3-4 category
                        ifelse(Nodule.lobes %in% c("3", "4"), "3-4", Nodule.lobes)))))%>%
#Making sure the branch coulmn is considered a factor
  mutate(Nodule.lobes = as.factor(Nodule.lobes))

#Extracting sample data from the sample-ID
cfu_df<-separate(cfu_df, Plant, into = c("Strain","Host","rep"), sep="\\.")

#Ensuring intended variables are numeric and factor as intended
cfu_df$rhz_released<-as.numeric(cfu_df$rhz_released)
cfu_df$Nodule.area<-as.numeric(cfu_df$Nodule.area)

#Removing 0.44.V.b.4 & 070.O.b.5  because of a sampling error making it a massive outlier in the nodule.area measure
Nodule.area.outliers<-c("044.V.b.4")
RhzReleased.outliers<-c("070.O.b.5")

cfu_df<-cfu_df %>%
#Removing size that was likely an error during processing
    filter(!(Ind.Label %in% RhzReleased.outliers))%>%
  filter(!(Ind.Label %in% Nodule.area.outliers))

#Removing any NA's introduced by adding a log transformation to the colony counts
cfu_df[is.infinite(log10(cfu_df$rhz_released)) | 
         is.na(log10(cfu_df$rhz_released)), ]
cfu_df <- subset(cfu_df, rhz_released > 0 & !is.na(rhz_released))

```

```{r Strain Reaction norm df (str.host_mean_df), collapse=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Another dataframe to create a reaction norm like plot
#Taking the mean across reps, strain, and variety to reduce noise
str.host_mean_df <- ss_df %>%
  group_by(Strain,Variety)%>%
  summarise(mean_prop_lobe=mean(prop_lobe,na.rm=TRUE), 
            mean_abvmass_g=mean(Shootweight_g,na.rm=TRUE),
            mean_belmass_g=mean(Rootweight_g, na.rm=TRUE),
            mean_biomass_g=mean(Biomass_g, na.rm=TRUE),
            mean_nod_num=mean(Nod_count, na.rm=TRUE),
            mean_agbiomass_g=mean(Shootweight_g, na.rm=TRUE))
```

#Statistics used 
##Rhizboial focused traits 
```{r Rhizobial Trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova examining Nodule area response predicted by branch #, host variety,and isolate ID
nodarea<-lm(log10(Nodule.area)~Nodule.lobes*Host*Strain+rep, data=cfu_df)
nodarea.results<-lm_run(nodarea, model_name="nodarea")
aov.nodarea<-anova(nodarea)
write.csv(aov.nodarea, file=file.path(table_dir, "Nodule-Area-response_anova.csv"))
aov.nodarea
aov.nodarea[["Sum Sq"]] / sum(aov.nodarea[["Sum Sq"]])

#Model examining Amount of viable rhizobia released focused on number of nodule branches as a predictor
rhz_released<-lm(log10(rhz_released)~Nodule.lobes*Host+as.factor(Strain)+rep, data=cfu_df)
rhz_released.results<-lm_run(rhz_released, "rhzreleased")
aov.rhzreleased<- anova(rhz_released)
write.csv(aov.rhzreleased, file=file.path(table_dir,"RhizobiaReleased-response_anova.csv"))
aov.rhzreleased
aov.rhzreleased[["Sum Sq"]] / sum(aov.rhzreleased[["Sum Sq"]])

# Model examining amount of rhizobia released focused on nodule surface area as a predictor 
rhztestarea<-lm(log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep, data=cfu_df)
rhztestarea.results<-lm_run(rhztestarea,"rhzarea")
aov.rhztestarea<- anova(rhztestarea)
write.csv(aov.rhztestarea, file=file.path(table_dir,"Nodulearea-by-surfacearea_anova.csv"))
aov.rhztestarea
aov.rhztestarea[["Sum Sq"]] / sum(aov.rhztestarea[["Sum Sq"]])
```

##Statistical Analysis of the Host plant traits
```{r host plant trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova is being ran on models associated with host traits 
#Modeling the mean nodule number predicted by proportion of branched nodules and host variety
nodnumber<-lm(mean_nod_num~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
nodnum.results<-lm_run(nodnumber, "nodnum")
aov.nodnum<-anova(nodnumber)
write.csv(aov.nodnum, file=file.path(table_dir,"Nodulenumber-response_anova.csv"))
aov.nodnum
aov.nodnum[["Sum Sq"]] / sum(aov.nodnum[["Sum Sq"]])

#Modeling the mean host plant biomass predicted by the mean proportion of branched nodules and host variety
biomass<-lm(mean_biomass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
biomass.results<-lm_run(biomass, "biomass")
aov.biomass<-anova(biomass)
write.csv(aov.biomass, file=file.path(table_dir,"Biomass-response_anova.csv"))
aov.biomass
aov.biomass[["Sum Sq"]] / sum(aov.biomass[["Sum Sq"]])

##proportion branches
propbr<- lm(prop_lobe ~ Variety*Strain+Rep,data=ss_df)
propbr.results<- lm_run(propbr,"meanbranch")
aov.propbr<-anova(propbr)
write.csv(aov.propbr, file=file.path(table_dir,"proplobe-response_anova.csv"))
aov.propbr
aov.propbr[["Sum Sq"]] / sum(aov.propbr[["Sum Sq"]])


#Modeling the mean host above and below ground biomass separately
##above ground
a.biomass<-lm(mean_abvmass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
a.biomass.results<-lm_run(a.biomass, "biomass")
aov.a.biomass<-anova(a.biomass)
write.csv(aov.a.biomass, file=file.path(table_dir,"Abovebround-biomass-response_anova.csv"))
aov.a.biomass
aov.a.biomass[["Sum Sq"]] / sum(aov.a.biomass[["Sum Sq"]])

##below ground
b.biomass<-lm(mean_belmass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
b.biomass.results<-lm_run(b.biomass, "Belowground biomass")
aov.b.biomass<-anova(b.biomass)
write.csv(aov.b.biomass, file=file.path(table_dir,"Belowground-biomass-response_anova.csv"))
aov.b.biomass
aov.b.biomass[["Sum Sq"]] / sum(aov.b.biomass[["Sum Sq"]])
```

#Figures
##Figure 2
```{r Figure 2, echo=FALSE, warning=FALSE, message=TRUE}
#Single plot exporting has been "#" out to make this knit, if an export is needed removing the "#" from the pdf() and dev.off() below.

#pdf(file=file.path(plot_dir,"Figure_2_surfacearea.pdf"), width=3, height=3)
surface_area_plot<- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=Nodule.area, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black", show.legend = FALSE)+
  labs(x="", y="Nodule Surface Area(cm2)", shape="Strain", color="Number of Nodule Branches")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  guides(size = "none",
         color = guide_legend(nrow = 1),
         shape = guide_legend(nrow = 1)) +
  plot_theme()
#dev.off()

#pdf(file=file.path(plot_dir,"Figure_2_released_rhizobia_branch.pdf"), width=3, height=3)
rhzreleasedbr_plot <- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=rhz_released, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black", show.legend = FALSE)+
  scale_y_log10()+
  labs(x= "Number of Nodule Branches", y="CFU/Nodule", shape="Strain", color="Number of Nodule Branches")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  guides(size = "none",
         color = guide_legend(nrow = 1),
         shape = guide_legend(nrow = 1)) +
  plot_theme()
#dev.off()

#pdf(file=file.path(plot_dir,"Figure_2_released_rhizobia_branch_ver5.pdf"), width=3, height=3)
rhzreleased_plot<-
  ggplot(cfu_df, aes(x = Nodule.area, y = rhz_released, color = Nodule.lobes)) +
  geom_point(size=3)+
  stat_smooth(method = "loess", span=10, se=FALSE, show.legend = FALSE) +
  scale_y_log10()+
scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  labs(x = "Nodule Surface Area(cm2)", y = "CFU/Nodule", shape="Strain", color="Number of Nodule Branches")+
    facet_wrap(~ Host, nrow = 3, labeller = labeller(Host = host_labels))+
  guides(shape="none", size="none")+ 
  plot_theme()
#dev.off()


#Grid format of Figure 2
Figure_2_plot<- (surface_area_plot / rhzreleasedbr_plot / rhzreleased_plot)+
  plot_layout(guides="collect", heights=c(1,1,1.75)) + 
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        legend.box = "vertical",
        plot.margin= unit(c(0.5, 0.5, 0.5, 0.5), "pt"),
        panel.spacing    = unit(1.5, "mm"),
        plot.tag = element_text(size = 26, face = "bold", hjust = 6, vjust= -2))
Figure_2_plot

svg(filename=file.path(plot_dir,"Figure_2.svg"), width=8, height=13)
print(Figure_2_plot)
dev.off()

```

##Figure 3

Maria: I think there are plots here that you don't really use in the paper that the one generated with the root weight. You don't need to keep the code if it is not important for the draft :). 
```{r Figure 3, echo=FALSE, warning=FALSE, message=TRUE}
#Plots generated for figure 3. Single plot exporting has been "#" out to make this knit, if an export is needed remove the "#" from the pdf() and dev.off() functions below.

#Mean Host Biomass Plot
#pdf(filename=file.path(plot_dir,"Figure_3_biomass.pdf"), width=3, height=3)
biomass_plot<-
ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_biomass_g, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(x="", y="Mean Host Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Host Shoot Biomass plot
#pdf(filename=file.path(plot_dir,"Figure_3_shoot_biomass.pdf"), width=3, height=3)
abiomass_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_abvmass_g, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(x="", y="Mean Host\nShoot Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

#Mean Host Root Biomass plot
#pdf(filename=file.path(plot_dir,"Figure_3_root_biomass.pdf"), width=3, height=3)
b.biomass_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_belmass_g, color=Variety))+ 
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(x="Mean Proportion of Branched Nodules ", y="Mean Host\nRoot Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

#Mean Nodule Count Plot
#pdf(filename=file.path(plot_dir,"Figure_3_nodulecount.pdf"), width=3, height=3)
nodulecount_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_nod_num, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(y="Mean Nodule Count", x="Mean Proportion of Branched Nodules",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Prop Nodules across Host Variety
#pdf(filename=file.path(plot_dir,"Figure_3_propbranch.pdf"), width=5, height=5)
proplobe_plot<-
  ggplot(str.host_mean_df, aes(x=Variety, y=mean_prop_lobe, group=Strain, color=Variety))+
  geom_line(color="gray",linewidth=1, alpha=0.3, show.legend = FALSE)+
  geom_jitter(alpha=0.7, width=0.25, size=3, show.legend = FALSE)+
  labs(x="Alfalfa Host Variety", y="Mean Proportion of\nBranched Nodules")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

# Generates the combined figure 3 plot
fi3panelA<- abiomass_plot + b.biomass_plot + plot_layout(ncol=2)

fig_3_plot <- fi3panelA  + nodulecount_plot + proplobe_plot +
  plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        legend.box = "vertical",
        plot.margin= unit(c(0.5, 0.5, 0.5, 0.5), "pt"),
        plot.tag = element_text(size = 26, face = "bold", hjust = 0.2, vjust = -4)) 
fig_3_plot



svg(filename=file.path(plot_dir,"Figure_3.svg"), width=10, height=10)
print(fig_3_plot)
dev.off()

```

#Supplemental

```{r Supplement tables}
#Building supplemental table of anova results from tested models
tab.1a <- data.frame(dataset="cfu_df", model="log10(Nodule.area)~Nodule.lobes*Host*Strain+rep",Term=row.names(aov.nodarea),
                     Df=aov.nodarea$Df,
                     Prop.Var=round(aov.nodarea$`Sum Sq`/sum(aov.nodarea$`Sum Sq`),3),
                     Fstat=round(aov.nodarea$`F value`,2),
                     Pvalue=aov.nodarea$`Pr(>F)`)
tab.1b <- data.frame(dataset="cfu_df", model="log10(rhz_released)~as.factor(Nodule.lobes)*Host+as.factor(Strain)+rep",Term=row.names(aov.rhzreleased),
                     Df=aov.rhzreleased$Df,
                     Prop.Var=round(aov.rhzreleased$`Sum Sq`/sum(aov.rhzreleased$`Sum Sq`),3),
                     Fstat=round(aov.rhzreleased$`F value`,2),
                     Pvalue=aov.rhzreleased$`Pr(>F)`)
tab.2a <- data.frame(dataset="str.host_mean_df", model="mean_nod_num~mean_prop_lobe*Variety+Strain",Term=row.names(aov.nodnum),
                     Df=aov.nodnum$Df,
                     Prop.Var=round(aov.nodnum$`Sum Sq`/sum(aov.nodnum$`Sum Sq`),3),
                     Fstat=round(aov.nodnum$`F value`,2),
                     Pvalue=aov.nodnum$`Pr(>F)`)
tab.2c <- data.frame(dataset="str.host_mean_df", model="mean_abovebiomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.a.biomass),
                     Df=aov.a.biomass$Df,
                     Prop.Var=round(aov.a.biomass$`Sum Sq`/sum(aov.a.biomass$`Sum Sq`),3),
                     Fstat=round(aov.a.biomass$`F value`,2),
                     Pvalue=aov.a.biomass$`Pr(>F)`)
tab.2z <- data.frame(dataset="str.host_mean_df", model="mean_belowbiomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.b.biomass),
                     Df=aov.b.biomass$Df,
                     Prop.Var=round(aov.b.biomass$`Sum Sq`/sum(aov.b.biomass$`Sum Sq`),3),
                     Fstat=round(aov.b.biomass$`F value`,2),
                     Pvalue=aov.b.biomass$`Pr(>F)`)
tab.s1 <- data.frame(dataset="rhztestarea", model="log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep",Term=row.names(aov.rhztestarea),
                     Df=aov.rhztestarea$Df,
                     Prop.Var=round(aov.rhztestarea$`Sum Sq`/sum(aov.rhztestarea$`Sum Sq`),3),
                     Fstat=round(aov.rhztestarea$`F value`,2),
                     Pvalue=aov.rhztestarea$`Pr(>F)`)
supplement_table <- rbind(tab.1a, tab.1b, tab.2a, tab.2c,tab.2z, tab.s1)
write.csv(supplement_table, file=file.path(table_dir, "Supplement_table.csv"))

```
