---
title: "Nodule branching, size, and symbiosis outcomes are shaped by natural genetic variation in rhizobial bacteria and alfalfa hosts "
author: "Elizabeth Leslie Paillan - elpailla@ncsu.edu & Liana Burghardt - liana.burghardt@psu.edu"
date: "2025-10-25"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

Nodule traits being examined: Proportion Lobed (prop_lobe= \# of lobed nodule/ total \# nodule sample) a lower proportion means fewer lobed nodules, and a higher proportion means larger number of nodules. Nodule Prevalence (nod_prev=total \# nodule sample), the term origniates from the avt_field experiement where I was dividing the number of nodules found by plant, that is not needed since its was one plant (mostly), for this experiment its just the total nodule number.

#Libraries Used

```{r Library setup, echo = TRUE, warning=FALSE}
library(tidyverse); packageVersion("tidyverse")
library(BSDA); packageVersion("BSDA")
library(patchwork); packageVersion("patchwork")
library(svglite); packageVersion("svglite")
library(performance); packageVersion("performance")
library(data.table); packageVersion("data.table")
set.seed(100)
```

```{r GGplot theme, include=FALSE}
#Below is the general theme used for all ggplot2 figures
plot_theme <- function(base_size = 18) {
  theme_bw(base_size = base_size) +
    theme(legend.title = element_text(face = "bold"),
      legend.text  = element_text(face = "bold"),
      axis.title   = element_text(face = "bold"),
      strip.text   = element_text(face = "bold"),
      panel.grid   = element_blank(),
      panel.background = element_blank(),
      axis.line    = element_line(linewidth = 1.05, color = "black"))
}

#Changing host variety names to their full names & add designated color scheme
host_labels <- c(O = "Onedia", S = "SW402", V = "Vernal")
host_colors <- c(O = "#471323", S = "#ADB9E3", V = "#FCB0B3")

#same thing while nodule branch number
nbranch_labels <- c("1", "2", "3-4", "5+")
nbranch_colors <- c("1" = "#273E47",
                    "2" = "#BAA5FF",
                    "3-4" = "#6E9887",
                    "5+" = "#A5402D")
nbranch_shapes <- c("1" = 16, "2" = 15, "3-4" = 17, "5+" = 18)

#and strain ID
strain_shapes <- modifyList(as.list(setNames(rep(16, length(unique(str.host_mean_df$Strain))),
                   unique(str.host_mean_df$Strain))),
                   list("001"=15, "012"=12, "044"=17, "059"=18, "070"=8, "077"=4, "112"=25, "117"=3))

strain_labels <- c("001"=1, "012"=2, "044"=3, "059"=4, "070"=5, "077"=6, "112"=7, "117"=8)
strain_shapes <- c("001"=15, "012"=24, "044"=17, "059"=18, "070"=8, "077"=4, "112"=25, "117"=3)



```

#Functions written for the analysis:

```{r Function: Anova on linear models -> pulling the summary and r2, echo=TRUE}
### This function runs a general anova, and spits out the summary() table and export it
lm_run<- function(model, model_name, save_path=getwd()) {
  anova_output<- anova(model)
  model_summary <- summary(model)
  results_list<- list()
  
  results_list[[paste0(model_name, "_anova")]] <- anova_output
  results_list[[paste0(model_name, "_summary")]] <- model_summary
  
  anova_df<- as.data.frame(anova_output)
  summary_df<- as.data.frame(model_summary$coefficients)
  
  return(results_list)
}
```

```{r Function: Detecting and Listing Outliers, }
#Maria: I think you might no longer be using the data coming out of detectview_outlier function. At least that is what I noticed with the dataframes/tables you created using them... maybe check and if you are not longer using them you can delete this chunk :)  

#This function is used to detect outliers, create a list of the identified outliers.
detectview_outlier <- function(data, variable) {
  outliers <- boxplot(data[[variable]], plot = FALSE)$out #Generates a barplot to ID outliers
  outlier_data <- data[data[[variable]] %in% outliers, ]
  
  outlier_plot <- ggplot(data, aes(x = 1, y = !!rlang::sym(variable))) + 
    geom_boxplot() + 
    geom_point(data = outlier_data, aes(x = 1, y = !!rlang::sym(variable)), color = "red") + #Makes the outlier more visible
    labs(x = "", y = variable, title = paste("Boxplot of", variable, "outliers"))
  
  return(list(plot = outlier_plot, outlier_data = outlier_data)) #Provides detail about the identified outliers
}
```

#Importing Data

```{r Data Import of Single Inoculation (ss_df), results = HIDE, warning = HIDE, collapse=TRUE}
#Import and tidying of Data
#setwd("~/GitHub/nodule_branching_morphology/Analysis/")
getwd()
#Setting up plot and directories to be easily found:
plot_dir<-("../Figures/") #Maria: I had to edit this to fit my directory, but if you use R.project and here() you will be able to make this more accessible for reviewers/readers. For Maria if I need to run this code again the path is /Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Figures/
table_dir<-("../Tables/"). # For Maria; If I need to run this code again the path is 

ss_df<-data.frame(fread("../data/MasterSpreadsheet_AVT.csv"))

# For Maria: If I need to run the code again the path is /Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/data/MasterSpreadsheet_AVT.csv"

# Removing erroneous columns in the original data frame, this includes collection notes
ss_df <- ss_df[, !colnames(ss_df) %in% "NOTES"]
#View(ss_df)
#nrow(ss_df) ## The number of rows in 1481

# Adding total nodule count and the proportion of branched nodules
ss_df<-ss_df %>%
  dplyr::select(Strain, Variety, Source_variety, Rep, Rootweight_g, Shootweight_g, Nod_lobe, Nod_unlobed, Label)%>% #Selecting columns relevant to the analysis
  mutate(Nod_count=Nod_lobe+Nod_unlobed)%>% #Counting total observed nodules
  mutate(prop_lobe=Nod_lobe/Nod_count) %>% #Creating the proportion of branched nodules to nodule count
  filter_at(vars(Nod_count:prop_lobe), all_vars(!is.infinite(.)))%>% #Checking for any infinite values
  mutate(Biomass_g=Rootweight_g+Shootweight_g)%>% #Counting total host mass biomass
  dplyr::filter(!Strain %in% c("MIX", "NIG", "MC2", "EXC", "UNI"))%>%
  drop_na() #Removing any NAs ## Maria: if you drop_na() in this part of the code, you will be eliminating samples for which you have relevant data, for example, you will eliminate 087_O_d, 110_O_b, 081_V_d, 080_O_d, etc (7 samples) for which we have shoot biomass data but we did not evaluate nodule number since so those cells have NAs in those nodule traits. There are not a lot of samples, but they could have been included in the phylogeny. Also, I don't know how important is Rootweight_g to be included as some of the traits to select() because when you do the drop_na(), it will drop 018_S_a that is a sample that although does not have rootweight it does have shoot biomass, and nodule numbers (It is the only relevant one I got to detect). I think it is important to drop the NAs when you are working with correlations, but that comes later in your pipeline. 

ss_df$Strain <- as.character(ss_df$Strain)

View(ss_df)

## Maria: I will save ss_df as a table to be able to open it in the other Rdm. 

write.csv(ss_df, "../Tables/ss_df.csv", row.names = FALSE) # For Maria next time to store this file: /Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/Tables/ss_df.csv



# Using the detect outlier function to isolate them for view if needed . # Maria: I don't see Biom_outlier or Lobe_outlier being used downstream so I don't think this part should be included in the code. Are you detecting these outliers to eliminate them from the downstream analysis?

Biom_outlier<-detectview_outlier(ss_df,"Biomass_g")
Biom_outlier ## Maria: It is giving me strain 020, 039, 057, 065 and 104. I am not really sure if you are filtering those values out?
Lobe_outlier<-detectview_outlier(ss_df,"prop_lobe")
Lobe_outlier ## Maria: It is giving me strain 009, 021, 039, 050, 080, 092, 110, 121


#Prop_lobe.outliers<-c("110_O_a")
#ss_df<-ss_df %>%
  dplyr::filter(!(Label %in% Prop_lobe.outliers))

# Separate data frame that is the mean of every Isolate-Host variety combination across all replicates
Mean_lobe<-ss_df %>% 
            group_by(Strain,Variety,Source_variety)%>% 
            dplyr::summarize(lobe_mean=mean(prop_lobe, na.rm=TRUE)) ##Maria: It doesn't seem like you use this dataframe in the rest of the pipeline so you might want to consider deleting it. This is very similar to the mean_prop_lobed in chunk near line 192. 
```

```{r Data Import of CFU experiment (cfu_df), message=FALSE, warning=FALSE, page.print=FALSE}
#Loading in the cfu dataset
cfu_df<-data.frame(fread("../data/AVT_CFU_count.csv"))
 ## For Maria, if I have to run the code again the path is /Users/mariaalejandragilpolo/Documents/GitHub/nodule_branching_morphology/data/AVT_CFU_count.csv
#View(cfu_df)
#Removing erroneous columns from the data frame like notes from the harvest, and renames one column to be easier to call into R

cfu_df <- cfu_df %>% 
  mutate(Comments=NULL)%>%
  mutate(dilution.counted=NULL)%>%  ## Maria: There is no need to remove this column because it is just raw data like the other counts :) Sincer you are starting with rhizobia.released.per.nodule.calculation, the raw values like dilution.counted and the raw counts are not really important because the calculation already happened in excel.
  rename(rhz_released = Lianas.rhizobia.released.per.nodule.calculation)%>% 
#Since we had a couple of nodules that had higher than 5 branches we combined them into a 5+ category
  mutate(Nodule.lobes = ifelse(Nodule.lobes == "more than 5", "5+", 
                        ifelse(Nodule.lobes == "More than 5", "5+",
                        ifelse(Nodule.lobes == "5", "5+",
#Few nodules had 3-4 branches so we combined them into a 3-4 category
                        ifelse(Nodule.lobes %in% c("3", "4"), "3-4", Nodule.lobes)))))%>%
#Making sure the branch coulmn is considered a factor
  mutate(Nodule.lobes = as.factor(Nodule.lobes))

#View(cfu_df)

#Extracting sample data from the sample-ID
cfu_df<-separate(cfu_df, Plant, into = c("Strain","Host","rep"), sep="\\.")

#Ensuring intended variables are numeric and factor as intended
cfu_df$rhz_released<-as.numeric(cfu_df$rhz_released)
cfu_df$Nodule.area<-as.numeric(cfu_df$Nodule.area)
View(cfu_df)
cfu_df$spot.avg <- as.numeric(gsub("[^0-9.]", "", cfu_df$spot.avg)) #Maria: I am not sure what this code is doing, but I don't think you need the spot.avg values since you are working with Lianas.rhizobia.released.per.nodule.calculation that already estimates the CFU from spot.avg and dilution. I think you can just remove this line. 
#View(cfu_df)


#Detecting outliers in the CFU data for view if needed (not used)
RhzReleased_outlier<-detectview_outlier(cfu_df,"rhz_released") #Maria: Similar to a previous note, I don't think you are using this information.
RhzReleased_outlier
Nodule.area_outlier<-detectview_outlier(cfu_df,"Nodule.area")
Nodule.area_outlier


#Removing 0.44.V.b.4  because of a sampling error making it a massive outlier in the nodule.area measure
Nodule.area.outliers<-c("044.V.b.4")
RhzReleased.outliers<-c("070.O.b.5")
View(cfu_df)
cfu_df<-cfu_df %>%
#Removing size that was likely an error during processing
    filter(!(Ind.Label %in% RhzReleased.outliers))%>%
  filter(!(Ind.Label %in% Nodule.area.outliers))

#Removing any NA's introduced by adding a log transformation to the colony counts
cfu_df[is.infinite(log10(cfu_df$rhz_released)) | 
         is.na(log10(cfu_df$rhz_released)), ]
cfu_df <- subset(cfu_df, rhz_released > 0 & !is.na(rhz_released))

```

```{r Strain Reaction norm df (str.host_mean_df), collapse=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Another dataframe to create a reaction norm like plot
#Taking the mean across reps, strain, and variety to reduce noise
str.host_mean_df <- ss_df %>%
  group_by(Strain,Variety)%>%
  summarise(mean_prop_lobe=mean(prop_lobe,na.rm=TRUE), 
            mean_abvmass_g=mean(Shootweight_g,na.rm=TRUE),
            mean_belmass_g=mean(Rootweight_g, na.rm=TRUE),
            mean_biomass_g=mean(Biomass_g, na.rm=TRUE),
            mean_nod_num=mean(Nod_count, na.rm=TRUE),
            mean_agbiomass_g=mean(Shootweight_g, na.rm=TRUE))
```

#Statistics used 
##Rhizboial focused traits 
```{r Rhizobial Trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova examining Nodule area response predicted by branch #, host variety,and isolate ID
nodarea<-lm(log10(Nodule.area)~Nodule.lobes*Host*Strain+rep, data=cfu_df)
nodarea.results<-lm_run(nodarea, model_name="nodarea")
aov.nodarea<-anova(nodarea)
write.csv(aov.nodarea, file=paste(table_dir,"Nodule-Area-response_anova.csv"))
aov.nodarea
aov.nodarea[["Sum Sq"]] / sum(aov.nodarea[["Sum Sq"]])

#Model examining Amount of viable rhizobia released focused on number of nodule branches as a predictor
rhz_released<-lm(log10(rhz_released)~Nodule.lobes*Host+as.factor(Strain)+rep, data=cfu_df)
rhz_released.results<-lm_run(rhz_released, "rhzreleased")
aov.rhzreleased<- anova(rhz_released)
write.csv(aov.rhzreleased, file=paste(table_dir,"RhizobiaReleased-response_anova.csv"))
aov.rhzreleased
aov.rhzreleased[["Sum Sq"]] / sum(aov.rhzreleased[["Sum Sq"]])

# Model examining amount of rhizobia released focused on nodule surface area as a predictor 
rhztestarea<-lm(log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep, data=cfu_df)
rhztestarea.results<-lm_run(rhztestarea,"rhzarea")
aov.rhztestarea<- anova(rhztestarea)
write.csv(aov.rhztestarea, file=paste(table_dir,"Nodulearea-by-surfacearea_anova.csv"))
aov.rhztestarea
aov.rhztestarea[["Sum Sq"]] / sum(aov.rhztestarea[["Sum Sq"]])
```

##Statistical Analysis of the Host plant traits
```{r host plant trait, echo=FALSE, warning=FALSE, message=TRUE}
#Anova is being ran on models associated with host traits 
#Modeling the mean nodule number predicted by proportion of branched nodules and host variety
nodnumber<-lm(mean_nod_num~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
nodnum.results<-lm_run(nodnumber, "nodnum")
aov.nodnum<-anova(nodnumber)
write.csv(aov.nodnum, file=paste(table_dir,"Nodulenumber-response_anova.csv"))
aov.nodnum
aov.nodnum[["Sum Sq"]] / sum(aov.nodnum[["Sum Sq"]])

#Modeling the mean host plant biomass predicted by the mean proportion of branched nodules and host variety
biomass<-lm(mean_biomass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
biomass.results<-lm_run(biomass, "biomass")
aov.biomass<-anova(biomass)
write.csv(aov.biomass, file=paste(table_dir,"Biomass-response_anova.csv"))
aov.biomass
aov.biomass[["Sum Sq"]] / sum(aov.biomass[["Sum Sq"]])

##proportion branches
propbr<- lm(prop_lobe ~ Variety*Strain+Rep,data=ss_df)
propbr.results<- lm_run(propbr,"meanbranch")
aov.propbr<-anova(propbr)
write.csv(aov.propbr, file=paste(table_dir,"proplobe-response_anova.csv"))
aov.propbr
aov.propbr[["Sum Sq"]] / sum(aov.propbr[["Sum Sq"]])


#Modeling the mean host above and below ground biomass separately
##above ground
a.biomass<-lm(mean_abvmass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
a.biomass.results<-lm_run(a.biomass, "biomass")
aov.a.biomass<-anova(a.biomass)
write.csv(aov.a.biomass, file=paste(table_dir,"Abovebround-biomass-response_anova.csv"))
aov.a.biomass
aov.a.biomass[["Sum Sq"]] / sum(aov.a.biomass[["Sum Sq"]])

##below ground
b.biomass<-lm(mean_belmass_g~mean_prop_lobe*Variety+Strain,data=str.host_mean_df)
b.biomass.results<-lm_run(b.biomass, "Belowground biomass")
aov.b.biomass<-anova(b.biomass)
write.csv(aov.b.biomass, file=paste(table_dir,"Belowground-biomass-response_anova.csv"))
aov.b.biomass
aov.b.biomass[["Sum Sq"]] / sum(aov.b.biomass[["Sum Sq"]])
```

#Figures
##Figure 2
```{r Figure 2, echo=FALSE, warning=FALSE, message=TRUE}
#Single plot exporting has been "#" out to make this knit, if an export is needed removing the "#" from the pdf() and dev.off() below.

pdf(file=file.path(plot_dir,"Figure_2_surfacearea.pdf"), width=3, height=3)
surface_area_plot<- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=Nodule.area, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black")+
  labs(x="", y="Nodule Surface Area(cm2)", shape="Strain", color="Number of\nNodule\nBranches")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  plot_theme()
dev.off()

pdf(file=file.path(plot_dir,"Figure_2_released_rhizobia_branch.pdf"), width=3, height=3)
rhzreleasedbr_plot <- 
  ggplot(cfu_df, aes(x=Nodule.lobes, y=rhz_released, color=Nodule.lobes, shape=Strain))+
  geom_jitter(aes(shape=Strain),width=0.2, size=3)+
  scale_size_manual(values=c(2.5))+
  geom_smooth(aes(group=Host),method = "lm", se=FALSE, color="black")+
  scale_y_log10()+
  labs(x= "Number of Nodule Branches", y="CFU/Nodule", shape="Strain", color="Host Variety")+
  scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  scale_shape_manual(values = strain_shapes, labels=strain_labels)+
  facet_wrap(~ Host, ncol = 3, labeller = labeller(Host = host_labels))+
  plot_theme()+
  theme(legend.position = "none")
  
dev.off()

pdf(file=file.path(plot_dir,"Figure_2_released_rhizobia_branch_ver5.pdf"), width=3, height=3)
rhzreleased_plot<-
  ggplot(cfu_df, aes(x = Nodule.area, y = rhz_released, color = Nodule.lobes)) +
  geom_point(size=3)+
  stat_smooth(method = "loess", span=10, se=FALSE) +
  scale_y_log10()+
scale_color_manual(values=nbranch_colors, labels=nbranch_labels)+
  labs(x = "Nodule Surface Area(cm2)", y = "CFU/Nodule")+
    facet_wrap(~ Host, nrow = 3, labeller = labeller(Host = host_labels))+
  plot_theme()+
  theme(legend.position = "none")

dev.off()


#Grid format of Figure 2
Figure_2_plot<- (surface_area_plot / rhzreleasedbr_plot / rhzreleased_plot)+
  plot_layout(guides="collect", heights=c(1,1,1.75)) + 
  theme(legend.position = "right")+
  plot_annotation(tag_levels = "A") &
  theme(plot.margin= unit(c(0.5, 0.5, 0.5, 0.5), "pt"),
        panel.spacing    = unit(1.5, "mm"),
        legend.box.margin = margin(0, 0, 0, 0, unit = "pt"),
        plot.tag = element_text(size = 26, face = "bold", hjust = 6, vjust= -2))
Figure_2_plot

svg(filename=file.path(plot_dir,"Figure_2.svg"), width=13, height=13)
print(Figure_2_plot)
dev.off()

pdf(file=file.path(plot_dir,"Figure_2_ver6.pdf"),width=10, height=14)
print(Figure_2_plot)
dev.off()
```

##Figure 3

Maria: I think there are plots here that you don't really use in the paper that the one generated with the root weight. You don't need to keep the code if it is not important for the draft :). 
```{r Figure 3, echo=FALSE, warning=FALSE, message=TRUE}
#Plots generated for figure 3. Single plot exporting has been "#" out to make this knit, if an export is needed remove the "#" from the pdf() and dev.off() functions below.

#Mean Host Biomass Plot
#pdf(filename=file.path(plot_dir,"Figure_3_biomass.pdf"), width=3, height=3)
biomass_plot<-
ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_biomass_g, color=Variety))+
  geom_point(alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  labs(x="", y="Mean Host Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Host Shoot Biomass plot
#pdf(filename=file.path(plot_dir,"Figure_3_shoot_biomass.pdf"), width=3, height=3)
abiomass_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_abvmass_g, color=Variety))+
  geom_point(aes(shape=Strain),alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(x="", y="Mean Host\nShoot Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

#Mean Host Root Biomass plot
#pdf(filename=file.path(plot_dir,"Figure_3_root_biomass.pdf"), width=3, height=3)
b.biomass_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_belmass_g, color=Variety))+ 
  geom_point(aes(shape=Strain) ,alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(x="Mean Proportion of Branched Nodules ", y="Mean Host\nRoot Biomass (g)",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

#Mean Nodule Count Plot
#pdf(filename=file.path(plot_dir,"Figure_3_nodulecount.pdf"), width=3, height=3)
nodulecount_plot<- 
  ggplot(str.host_mean_df, aes(x=mean_prop_lobe, y=mean_nod_num, color=Variety))+
  geom_point(aes(shape=Strain), alpha=0.6, size=3)+
  stat_smooth(method="lm",se = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(y="Mean Nodule Count", x="Mean Proportion of Branched Nodules",
       color= "Host Variety")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
  plot_theme()
#dev.off()

#Mean Prop Nodules across Host Variety
#pdf(filename=file.path(plot_dir,"Figure_3_propbranch.pdf"), width=5, height=5)
proplobe_plot<-
  ggplot(str.host_mean_df, aes(x=Variety, y=mean_prop_lobe, group=Strain, color=Variety))+
  geom_line(color="gray",linewidth=1, alpha=0.3, show.legend = FALSE)+
  geom_jitter(aes(shape=Strain),alpha=0.7, width=0.25, size=3, show.legend = FALSE)+
  scale_shape_manual(values=unlist(shape_vals), breaks = c("001", "012", "044", "059", "070", "077", "112", "117"), labels=strain_labels)+
  labs(x="Alfalfa Host Variety", y="Mean Proportion of\nBranched Nodules")+
  scale_color_manual(values =c("O"="#471323", "S"="#ADB9E3", "V"="#FCB0B3"))+
    plot_theme()
#dev.off()

# Generates the combined figure 3 plot
fi3panelA<- abiomass_plot + b.biomass_plot + plot_layout(ncol=2)

fig_3_plot <- fi3panelA  + nodulecount_plot + proplobe_plot +
  plot_layout(guides = "collect") + plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom", plot.tag = element_text(size = 26, face = "bold", hjust = 0.2, vjust = -4)) 


svg(filename=file.path(plot_dir,"Figure_3.svg"), width=13, height=13)
print(fig_3_plot)
dev.off()

#Doesnt have the fixed axes label and "other" label in legend
pdf(file=file.path(plot_dir,"Figure_3.pdf"), width=10, height=10)
print(fig_3_plot)
dev.off()
```

#Supplemental

```{r Supplement tables}
#Building supplemental table of anova results from tested models
tab.1a <- data.frame(dataset="cfu_df", model="log10(Nodule.area)~Nodule.lobes*Host*Strain+rep",Term=row.names(aov.nodarea),
                     Df=aov.nodarea$Df,
                     Prop.Var=round(aov.nodarea$`Sum Sq`/sum(aov.nodarea$`Sum Sq`),3),
                     Fstat=round(aov.nodarea$`F value`,2),
                     Pvalue=aov.nodarea$`Pr(>F)`)
tab.1b <- data.frame(dataset="cfu_df", model="log10(rhz_released)~as.factor(Nodule.lobes)*Host+as.factor(Strain)+rep",Term=row.names(aov.rhzreleased),
                     Df=aov.rhzreleased$Df,
                     Prop.Var=round(aov.rhzreleased$`Sum Sq`/sum(aov.rhzreleased$`Sum Sq`),3),
                     Fstat=round(aov.rhzreleased$`F value`,2),
                     Pvalue=aov.rhzreleased$`Pr(>F)`)
tab.2a <- data.frame(dataset="str.host_mean_df", model="mean_nod_num~mean_prop_lobe*Variety+Strain",Term=row.names(aov.nodnum),
                     Df=aov.nodnum$Df,
                     Prop.Var=round(aov.nodnum$`Sum Sq`/sum(aov.nodnum$`Sum Sq`),3),
                     Fstat=round(aov.nodnum$`F value`,2),
                     Pvalue=aov.nodnum$`Pr(>F)`)
tab.2c <- data.frame(dataset="str.host_mean_df", model="mean_abovebiomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.a.biomass),
                     Df=aov.a.biomass$Df,
                     Prop.Var=round(aov.a.biomass$`Sum Sq`/sum(aov.a.biomass$`Sum Sq`),3),
                     Fstat=round(aov.a.biomass$`F value`,2),
                     Pvalue=aov.a.biomass$`Pr(>F)`)
tab.2z <- data.frame(dataset="str.host_mean_df", model="mean_belowbiomass_g~mean_prop_lobe*Variety+Strain",Term=row.names(aov.b.biomass),
                     Df=aov.b.biomass$Df,
                     Prop.Var=round(aov.b.biomass$`Sum Sq`/sum(aov.b.biomass$`Sum Sq`),3),
                     Fstat=round(aov.b.biomass$`F value`,2),
                     Pvalue=aov.b.biomass$`Pr(>F)`)
tab.s1 <- data.frame(dataset="rhztestarea", model="log10(rhz_released)~Nodule.area*Nodule.lobes+Host+Strain+rep",Term=row.names(aov.rhztestarea),
                     Df=aov.rhztestarea$Df,
                     Prop.Var=round(aov.rhztestarea$`Sum Sq`/sum(aov.rhztestarea$`Sum Sq`),3),
                     Fstat=round(aov.rhztestarea$`F value`,2),
                     Pvalue=aov.rhztestarea$`Pr(>F)`)
supplement_table <- rbind(tab.1a, tab.1b, tab.2a, tab.2c,tab.2z, tab.s1)
write.csv(supplement_table, file=paste0(table_dir, "Supplement_table1.csv"))

```
